### 2019-09-26 Old/body_acm.tex

\section{Introduction}

An evolving consumer culture is has led wireless internet service providers
(ISPs) to rethink their service plans. Mobile data usage is quickly outpacing voice and SMS in
wireless network, and the trend is only expected to increase with
multi-device ownership. Declining revenue has caused ISPs search for sources
of new revenue in the changing market. Thus the introduction of the shared data plan \cite{att}. Using an account
service, users are able to keep track of data usage in real time across all
their devices. 
The shared data service plan requires that users hold an \emph{a priori}
knowledge of demand.
We address several topics: data as a product in the real-monetary market, and
data a network resource in a wireless topology.

Many new services are found exclusively on mobile devices; older softwares
are moving from (wired) grid-based to node-based communication. Software-defined
networking (SDN) addresses the new environment of wireless communication
devices, allowing for a programmable network architecture. 
The account services that manage wireless shared data plans decentralize
network management, and mobility becomes a factor in SDN design.
Individual mobile devices provide flexibility, and may make decisions regarding
local network infrastructure. There is a clear need for algorithms designed for
optimization in this space. In many cases, the direct communication between
mobile devices allows for a simple mutation of classic optimization models.
Auctions are key in SDN for the fair allocation of resources. For this work, we
focus on mobile data, an infinitely divisible and distributable quantity.
Mobile data represents online data accessed using the WISP network, and as
representation of future network usage, we are able remove restrictions imposed
by the physical layer. 
In \cite{lazar}, Lazar and Semret introduced the Distributed Progressive Second Price (PSP)
Mechanism for bandwidth allocation.
auction mechanisms that are (1) easily distributed, and (2) allocate an infinitely
divisible resource.
An auction mechanism is defined
as distributed when the allocations at any element depend only on
local state, no single entity holds a global market knowledge. We consider the
multi-auction; each auctioneer is a user selling data to its peers.

The model for data exchange was recently adopted by China Mobile Hong Kong
(CMHK), who released a platform, 2cm (2nd exchange market) creating a secondary
market where users can buy and sell data from each other.
CMHK owns and moderates 2cm, centrally computing allocations of mobile data
based on bids submitted to the platform.
Flexible data-sharing
plans are similar to the CMHK market, a limited number of devices may
share a single data plan. A shared-data plan, however limited, offers better economy
by creating primary users with a service package with cellular and a lot of data, and
limited number of secondary users that are using only data \cite{att}.
The secondary market allows for primary and secondary users to freely correspond,
without the restriction of a static primary-secondary user association.

Users on the shared data plan given by \cite{att} do not ``buy" or ``sell"
data, however, we may easily augment the model to include a price function,
which may be virtual, creating a secondary market.
In order to demonstrate potential of expanding to other $\emph{a priori}$ use
models, we give a simple example. Consider Alice and Bob, who met through an online service connecting neighborly
individuals closest to each other, in fact, close enough to be within wireless
range. Alice is going on vacation, and offers to transfer her wireless signal to
Bob. As they have the same provider, Bob agrees, and is able to use Alice's
bandwidth, which he finds is useful during peak network
hours. He pays a discounted rate. We note that the additional bandwith may be
used by another device, or even a 5G-capable mobile device.

The secondary market provides a unique opportunity for social equilibrium, as it
allows users to share data without sharing the same data plan, a restriction in
most ISPs, such as \cite{att}. We reason that a secondary market effectively creates a competitive secondary market, and contributes to the
dynamics of a free-market economy. Market competition is a desirable quality in
free markets, and is encouraged, particularly in wireless and data services. In
fact, California Legistature has recently passed laws promoting
competition and enforcing fair practice of ISPs \cite{cabill}.
Laws such as \cite{fedbill} exist to
regulate ISPs as they have historically come close to monopolizing regional
markets, leading to consumer abuse.
The global view of privacy has not been addressed, as the data exchange model is still
in its experimental phase.
Within the secondary market, bid privacy is a concern for two reasons: (1) Buyers are reluctant to
reveal their true valuations, as sellers may use these values to discriminate
against specific buyers. (2) Buyers doubt an auction's outcome, as they do
not pay what they bid, e.g. the auctioneer might create a fake
second highest bid slightly below the highest bid in order to increase his revenue
In general, the buyer does not trust the auctioneer, and the economy does not
trust the ISP.
We therefore determine that our mechanism must be (3) globally and locally privacy-preserving.

The market
topology and the user strategy are organically determined by the impact of user
behavior on market dynamics, and so determines a minimally optimal objective
representing user valuation globally, and so fufills an additional property of
economy over time and space. To the best of our knowledge, this is the first
work to provide a comprehensive derivation of a truthful mechanism that is self-contained within a dynamic market
topology.
To the best of our knowledge, this is the first work to provide a comprehensive
derivation of a truthful mechanism that is self-contained within a dynamic market
topology.

In classic mechanism design, with multiple user types, there is no single
way to design the transformation from the direct revelation mechanism to its
corresponding computational design. We apply a modifier to the PSP mechanism in
order to mutate the strategy space, following dynamic user
correspendence. As in \cite{lazar}, we take the direct
approach by guessing the right modifier, and context, such that we have the
desired result by composition with the PSP rules. 
As in \cite{lazar}, the incentive for a
user to truthfully reveal its type is built into the user strategies. Then,
local equilibria follow as a result of incentive
compatibility characterizing best strategy moves.
We claim that our formulation holds the desired PSP qualities,
%converging at a rate of $\mathcal{O}(\sum_{i\in\mcI}(\sum_{j\in\mcI}) \theta_i /
%\epsilon I!)$.
%We note that the convergence time is 
and that our strategy profiles are natural; they depend on the ratio of
supply and demand, or the ratio of buyers to sellers.

We focus on providing users with an incentive framework, and
so rational users choose a collaborative exchage. 
The user strategies are organic in that they are natural, or
induced by the dynamic market itself. In other words, adhering to the second-price rule,
where price is derived from autonomous demand, we have a strategic
progressive auction, and a multi-objective equilibria.
This is the (built-in) transformation from the
direct-revelation mechanism to the desired message space, Then, in the limit of
the data-model, a user reveals its valuation of a quantitiy of data-resource over the whole range of possible demands.

We describe our auction mechanism as a
pure-strategy progressive game with incomplete, but perfect information.

\section{Related Work}

Progressive second price auctions are used for optimal allocation in a
variety of scenarios, and for different reasons. Different definitions of social
welfare define different strategies. Typical goals of optimization are the
maximization of revenue, and optimal allocation. Other papers' focus, taken from
auction theory, optimize sellers' reserve price, or market price. Results derived from game theory
focus on player strategy, as in this work. In \cite{qpsp}, user strategy gives a ``quantized"
version of PSP, improving the rate of convergence of the game. Modifications to
the mechanism that result in improved convergence also appear in \cite{burst},
which relies on a global approximation function of demand.
Another mechanism derived from game theory \cite{diffserve}, derives
optimal strategies for buyers and brokers, and further shows the existence
of networkwide market equilibria by provisioning a dynamical system. The stability
of the game may be implied by various equilibria in distributed systems.

Allowing a user preference to, loosely, represent a policy, we may interpret
a user preference from the data exchange market as a meta-policy. Then,  plans allow users to set their own policies, and rely existing
framework for to implement their preferences. Several studies address the
security of policy-based network management.
Trusted management systems based on the
Common Informtion Model (CIM) focus on policy-based management, namely,
the ``Policy-Maker" toolkit. 
In general, the translation of policy-based management systems to SDN focuses on combining
the simplicity of policy-based implementation with the flexibility of SDN, as in the
meta-policy system, CIM-SDN \cite{meta}.

In the general context of distributed and decentralized allocation of
resources, a variety of equilibria exist for heterogeous and homogenous services
with the similar condition of truthfulness. In all cases, the PSP constraints are
sufficient to attain the desirable property of truthfulness through incentive
compatibility, as the pricing mechanism upholds the exclusion-compensation
principle.

\section{The Problem Model}

\subsection{The Secondary Market}

We construct the model for a PSP data auction for mobile users
particpating in secondary mobile data exchange market.
Let the set of all wireless users to be labeled by
the index set $\mcI = \lbrace 1, \cdots, I\rbrace$. In our current formulation, we do not allow a seller to host multiple
auctions, thus we may assume that data is a unary resource belonging to the seller, and
identify each local auction with the index of the seller
$j\in\mcI$. The bid profiles of the users are given as,
$s \equiv
[s_i^j] \in \mcI \times \mcI$,  
We have the strategy space for buyer $i$ as all possible bids at all auctions:
$S_i=\Pi_{j\in\mcI} S_i^j$, and $S_{-i} =\Pi_{j\in\mcI} \big( \Pi_{k\ne
i\in\mcI} S_k^j\big)$ as the
associated opponent profiles.

The grid of bid profiles, $s$, represents the state of distributed PSP auction
mechanism in the secondary market. 
We emphasize that we allow the grid $s$ to represent the bids of all buyers and
sellers. In general, we will not reference the full grid. In order to emphasize
that a bid belongs to a seller, we use the notation $\s_i^j$. We
will also empasize the \emph{context} of the bid to indicate the user type.
To further clarify our analysis, we adopt the notational conventions: a
seller's profile is denoted by $\s^j = [s_i^j]_{i\in\mcI}$, and $s_i =
[s_i^j]_{j\in\mcI}$ denotes a buyer's profile, where $s_{-i} \equiv [s_1^j , \cdots , s_{i-1}^j , s_{i+1}^j , \cdots
, s_I^j]_{j\in\mcI}$ as the profile of user $i$'s opponents.
Furthermore, noting that this is a simplification for ease of notation, we let
$D^j = \sum_{i\in\mcI} d_i^j$ be the total amount of data $j$ has to sell, and
$D_i = \sum_{j\in\mcI} d_i^j$ represent the total amount of data desired by buyer $i$.

We assume a public platform, published by the ISP, that allows sellers to advertise their
auctions, and that buyers may submit bids directly to sellers over the wireless
network. 
The ISP is included
by introducing a mutation of user type, and represent
the ISP as a blind, deaf user $\kappa$, who does not participate
in any auctions, but nonetheless holds the power to create a monopoly
.
We assume that buyers and sellers are
separated (a seller does not also buy data and vice versa). 
In general, we denote a buyer's identity $i \in \mcI$, and a seller as
$j\in\mcI$.
Suppose $i$ is buying from $j$. We assume that a user's budget is 
sufficient, as the alternative would be to pay a higher price to the ISP. The bid is represented by $s_i^j = (d_i^j, p_i^j)$, 
meaning $i$ would like to buy from $j$ a quantity $d^j_i$ and is willing to pay
a unit price $p^j_i$. 
Seller $j$'s local auction begins at time $t>0$. The seller takes reponsibility
to send opponent bid profiles $s_{-i}$ to each buyer that joins the auction,
those buyers where $s_i^j > 0 \in s$, as well its own bid, $\s_i^j = (d_i^j,
p_i^j)$, offfering quantity $d_i^j \in d^j = [d_i^j]_{i\in\mcI}$, 
with reserve unit price $p_i^j \in p^j = [p_i^j]_{i\in\mcI}$. 
Naturally, in a live auction, if a buyer does
not submit a bid to a seller, then this implies $s_i^j=\s_i^j = 0$. A buyer that does not submit a bid will not recieve opponent profiles from seller
$j$. We additionally determine that a user who does not submit a bid is holding
to the previous bid, either zero or nonzero. We note that buyers are
consistently referenced using the index $i$ as a subscript, and sellers using
the index $j$ as a superscript, as in \cite{semret}.

We will examine the role of buyers, who are able to directly
influence global market dynamics, and assume that the sellers take a reactionary role.
Each buyer $i$ will have information from each seller $j$, as well as opponent
profiles $s_{-i}$, from each auction in which it is participating.
In the extreme case, where $i$ submits bids to all auctions $j\in\mcI$, buyer
$i$ gains access all buyer profiles, $[s_1, \cdots, s_I]$. However, 
sellers can only gain information about the market grid by observing buyer
behavior in their local auction.
Finally, we define the seller pool for buyer $i\in\mcI$: 
% MIN SET
$$
    \mcI_i(n) =\argmax_{\mcI' \subset \mcI, \vert\mcI'\vert =
n}\sum_{j\in\mcI'} D^j,
$$
and similarly, for a seller $j\in\mcI$, we define the set of participating buyers:
$$
    \mcI^j(m) =\argmax_{{\mcI}' \subset \mcI, \vert{\mcI}'\vert =
m}\sum_{i\in{\mcI}'} p_i^j,
$$
where $m, n\in\mcI$.

\subsection{The Data Market Problem}

We aim to design a distributed PSP auction, operating within a strategic framework that
determines the bidding behavior of users in a wireless network.
The auction design must must meet a certain set of known
criteria: (1) \emph{truthfulness}, (2) \emph{individual rationality/
selfishness}, (3) \emph{social welfare maximization}, and (4)\emph{the winning
bid is persistent private information}. For the secondary data exchange
market, we determine that the
strategy space must meet additional criteria: (5) \emph{privacy and
independence from the ISP}, (6) \emph{locally fair division}, and (7)
\emph{minimize crossover in buyer/ seller pools}.
Thus, we propose a mechanism that promotes social welfare. 
Thus our motivation to adapt the PSP auction. The PSP auction given in
\cite{lazar} is comprised of a set of simple and symmetric rules that closely
follow market theory, and as it is distributed we can require privacy.

We first address the clear need for privacy in the secondary data exchange market. In \cite{zheng},
it is assumed that the ISP interferes in 2cm (market) dynamics, and will maximize the
gap between supply and demand in each transaction, exacting the difference as
revenue. We notice that this market behavior is suspiciously monopolistic, as a
single entity, the ISP, holds a global market power. It is further claimed in
\cite{zheng} that user bids are truthful as they are guaranteed to receive
their bid. We argue that this model represents an ``unwitting" buyer and an
equally uninformed seller, as they have no intution of fair market value.
Thus, in the interest of social good, we aim to provide a method to arrest
anti-competitive conduct by ISPs. Then, we may trust the ISP. Naturally,
rational users will only purchase data if the rewards outweigh the risk, it follows that
the secondary auction should be trusted as well. 
We argue that as the mechanism is also privacy-preserving, that both the users and the ISP benefit.
An ISP that is trusted will end up with a loyal customer base, taking profit
from to one that is revenue maximizing, assuming that competion laws will
prevent a monopoly.

We define a value to be private if any coalition is incapable of learning any information
besides what can be inferred from the shared computation and the coalition’s
inputs. We describe the process as given in \cite{felix}. In general, a distributed private
computation, where buyer $i$ is part of a coalition comprising auction $j$, is as follows:

Denoting $m_{-i} = [(s_i^j, r_i), m_1, \cdots, m_n]_{k\ne i\in \mcI}$, buyer sends a
message to each of its opponents, where $s_i^j$ is $i$'s bid, $r_i$
is an independent random value, and $m_1, \cdots, m_n$ the messages $i$ has
received so far.
Then, all buyers are able to confirm the winning bid $s_{\bar{i}}^*$.

The
private computation process requires that buyers connect with each other, and
so reveal their identities to the coalition. It was proven in \cite{felix} that full privacy is not possible in a second
price auction, even if we allow partial relevation and weak coalitions. Yet, we
claim that our mechanism is trustworthy:

\subsubsection{Buyers are anonymous}
In our secondary market, we have that any local auction is anonymous by definition, as a permutation of
the valuations resutls in a permutation of allocations and prices, equivalently,
exchanging the bids of two losing buyers does not change the auction's result.
Formally,
\definition{(Anonymous auction)} \cite{felix}
Given an auction $j$ and buyers $i\in\mcI$, a protocol for computing
$\max\lbrace i \in\mcI : p_i^j \ge p_k^j \ \forall \ k\in\mcI\rbrace$ if for
all coalitions $T\subset \mcI$, any pair of inputs $x = [s_1^j, \cdots,
s_I^j], \xi$, so that $\xi$ is a permutation of $x$, $\forall \ i \in T : x_i =
\xi_i$, and $\max( )$, and any choice of random inputs $\lbrace
r_i\rbrace_{i\in T}$. Let $\bar{T} =   T \times \mcI \setminus T$, 
$$
    Pr\big([x, \lbrace r_i\rbrace_{i\in T}]_{x\in}
    \big\vert
    \lbrace r_i\rbrace_{i\in T}\big) = Pr\big([\xi, \lbrace r_i\rbrace_{i\in T}]_{\xi\in T
    \times \mcI \setminus T}\big\vert
    \lbrace r_i\rbrace_{i\in T}\big)),
$$
which states that any two inputs, the messages seen by coalition $T$
are indentically distributed.

\subsubsection{The winning bid is persistent private information.}
We claim that a buyer's trust in a local auction is fufilled when the outcome of
the auction is guaranteed to be correct, and if the winner's identity is secret.
For each local auction, we define a coalition of the
participating buyers. The winning bidder is \emph{privately} chosen by
distributing the computation of the winner to the local coalition. The
computation is based on the homomorphicity of secret sharing. We present the
Lemma in its general form,
{
\lemma{(Benaloh 1987)}\label{modp} \ 
    $f(x_1, x_2, \cdots, x_n) = \sum_{i=1}^n x_i \mod p$ is privately
    computable.
}
Thus, it is possible to privately compute,
\begin{equation}\label{winner}
    \omega = \max([p_i^j]_{i\in\mcI^j}) = (p_1^j, p_2^j, \cdots, p_n^j, \argmax(p^j).
\end{equation}
We have the following procedure determining the winner of auction $j$ for
some fixed time in the bid progression, where $D^j > 0$.
% PRIVACY
\begin{center}
\begin{algorithm}
\caption{(Max bid private computation)}
\begin{algorithmic}[1]\label{maxbidcomp}
\State $\omega\gets 1, e\gets 0$
\While{$e \le 1$}
\For{$i \in \mcI^j$}
\If{$p_i^j \le \omega$}
\State $p_i^j \gets 1 $
\Else
\State $p_i^j \gets 0$
\EndIf
\EndFor
\State $e = \sum_{i\in\mcI^j} p_i^j \mod (n+1)\quad$ (Lemma \ref{modp})
\For{$i \in \mcI^j$}
\If{$p_i^j \ge e$}
\Return $i \ (winner)$
\EndIf
\EndFor
\EndWhile\\
\end{algorithmic}
\end{algorithm}
\end{center}

The winning buyer then leaves the auction, and so we have that the privacy of
the winning bid is persistent. We note that it is possible for a winner to
anonymously rejoin an auction, however this does not alter our result.

\subsubsection{Privacy and independence from the ISP.}
In our model, privacy is integrated into the mechanism. Our design enforces
privacy by hiding the reserve price of the sellers. At time
$t=0$, a seller $j$ entering the market will
submit bid $\s_\kappa^j =(D^j, \epsilon)$ to the public data exchange
platform, and so the initial bid $\s_\kappa^j$, is public knowledge. 
The auction
begins at time $t>0$, and at $t=0$, $j$ will
initializing its reserve price by executing a single bid iteration.
Sellers do not update pricing information with the ISP, thereby hiding
its local market price in the data-exchange market. As the ISP has limited
information from its ``competitor", it is unable to sabatoge prices derived
from fair market competition.
Thus, we claim that our model supports and protects the secondary market, allowing it
to be in direct competition with its parent ISP, and so contributes to the
regulation of ISPs \cite{fedbill} and supports a regional free-market economy
with respect to wireless data \cite{cabill}.
We will assume that the cost of participating in the secondary market is
absorbed by the bid fee, which could represent data used in submitting bids, or
a fee charged per unit of data, or a flat rate charged at the completion of the
purchase. We do not model ISP revenue, but assume it may be extracted from the
bid fee at $t=0$.

We now proceed to formally define the PSP auction, which determines the actions 
buyers and sellers in the secondary market, and which we will denote the \emph{data}
PSP rules. 
We define an \textbf{opt-out function}, $\g_i$, associated with a buyer $i$ as
part of its type. Buyer $i$, when determining how to aquire a possible allocation $a$,
will determine its bid quanities by,
\begin{equation}\label{opt-out}
    \g_i(a) = [\g_i^j(a)]_{j\in\mcI}.
\end{equation}
In a general sense, $\g_i$ applies our user strategy to the PSP rules.

%\subsection{PSP for Data-Exchange}
% MECHANISM
\subsubsection{The Mechanism}\label{datamechanism}
The rules presented here incorporate of the opt-out function with the
mechanism as in \cite{lazar}, which we note greatly simplifies our
analysis. 
The market price function (MPF) for a buyer in the secondary market
can be described as follows:
% NEW MARKET PRICE
\begin{align}\label{dataprice}
\begin{split}
    &\bar{P}_i(z, s_{-i}) =\displaystyle\sum_{j\in\mcI} \ \g_i^j \circ P_i^j(z_i^j,
s_{-i}^j) \\
    &= \sum_{j\in\mcI}\bigg(\inf\bigg\lbrace y\ge 0 : 
    {D_i^j}(y,s_{-i}^j) \ge \g_i^j(z) \bigg\rbrace \bigg),\\
\end{split}
\end{align}
and is interpreted as the aggragate of minimum prices that buyer $i$ bids in
order to obtain
data amount $z$ given opponent profile $s_{-i}$. We note that in the following
analysis
the total minimum price for the buyer cannot be an aggragation of the
\emph{individual} prices of the buyers, as it is possible that the reserve prices
of the sellers may vary.\\
\textbf{Remark:} We further note that except at points of discontinuity, from
Lemma \ref{uservaluation} we have that $P_i^j(z) = f_i(z)$. \\
\iffalse
(THE ABOVE IS GOOD, BUT DOESN'T FIT MY CONSTRUCTION, CHANGE TO BELOW?)
The market price function (MPF) for a buyer in the secondary market
is determined per (\ref{buyerinversedemand}), and is defined as,
% NEW MARKET PRICE
\begin{align}\label{dataprice}
\begin{split}
    &\bar{P}_i(z, s_{-i}) =\displaystyle\sum_{j\in\mcI} \ \g_i^j \circ P_i^j(z_i^j,
s_{-i}^j) \\
    &= \sum_{j\in\mcI}\bigg(\inf\bigg\lbrace y\ge 0 : 
    {D_i^j}(y,s_{-i}^j) \ge \g_i^j(z), \ \forall \ j\in\mcI \bigg\rbrace \bigg),\\
\end{split}
\end{align}
and is interpreted as the price that buyer $i$ bids in
order to obtain
data amount $z$ given opponent profile $s_{-i}$.
The sellers pricing function is according to (\ref{sellerinversedemand}), 
\begin{align}\label{dataprice}
\begin{split}
    &\bar{P}^j(z, s_{-i}) =\displaystyle\sum_{i\in\mcI} \ \g_i^j \circ P_i^j(z_i^j,
s_{-i}^j) \\
    &= \sum_{j\in\mcI}\bigg(\sup\bigg\lbrace y\ge 0 : 
    {D_i^j}(y,s_{-i}^j) \ge \g_i^j(z), \ \forall \ i\in\mcI \bigg\rbrace \bigg).\\
\end{split}
\end{align}
We note that
the total price cannot be an aggragation of the
\emph{individual} bid prices as it is possible that the reserve prices
of the individual sellers may vary, which contradicts
(\ref{buyerinversedemand}) and (\ref{sellerinversedemand}).\\
\textbf{Remark:} We further note that except at points of discontinuity, from
Lemma \ref{uservaluation} we have that $P_i^j(z) = f_i(z)$. \\
%To simplify the notation, we define $z_i^j = d^j / \vs_i^j$ opt-out
%buyer $i$'s bid quantity. 
\fi
The maximum available quantity of data in auction
$j$ at unit price $y$ given $s_{-i}^j$ is: 
% NEW ALLOCATION RULE
\begin{align}
\begin{split}\label{datacomposed}
    \bar{D}_i^j(y,s_{-i}^j) = \g_i^j \circ D_i^j( y,s_{-i}^j)
    = \bigg\lbrack D^j - \sum_{p_k^j> y} \g_k^j(a)\bigg\rbrack^+.
\end{split}
\end{align}
It follows from the upper-semicontinuity of $D_i^j$ that for $s_{-i}^j$ fixed, $\forall \ y,z \ge 0$,
\begin{equation}\label{dtop}
    \g_i^j(z) \le \g_i^j \circ D_i^j(y,s_{-i}^j) \Leftrightarrow y \ge \g_i^j
\circ P_i^j(z,s_{-i}^j).
\end{equation}
% DATA ALLOCATION RULE
The resulting data allocation rule is a function of the local market
interactions between buyers and sellers over all local auctions, as is composed with $i$'s opt-out value, so that for each $i \in \mcI$, the
allocation from auction $j$ is,
% and so for profile $s_i$, for any $i \in \mcI$,
\begin{align}\label{dataallocation}
\begin{split}
    \bar{a}_i^j(&s) = \g_i^j \circ a_i^j(s) \\
    &= \min\bigg\lbrace \g_i^j(a), 
    \frac{\g_i^j(a)}{\displaystyle\sum_{p_k^j= p_i^j}\g_k^j(a)} 
D_i^j(p_i^j,s_{-i}^j)\bigg\rbrace,
\end{split}
\end{align}
noting that for the full allocation from all auctions we may simply aggregate
over the seller pool. \\
\textbf{Remark:} The bid quantity $\g_i^j(a)$ and the allocation $\bar{a}_i^j$
are complementary. In fact, the buyer strategy is the first term in the
minimum, the second term being owned by the seller. \\
Finally we must have that the cost to the buyer adheres to the second price
rule for each local auction, with total cost to buyer $i$,
% DATA COST
\begin{align}\label{datacost}
\begin{split}
    \bar{c}_i(s) &= \displaystyle\sum_{j\in\mcI} 
    p_i^j \bigg(\bar{a}_i^j(0; s_{-i}^j)
    -\bar{a}_i^j(s_i^j;s_{-i}^j)\bigg).
\end{split}
\end{align}
\textbf{Remark:} The cost to buyer $i$ adds up the willingness
of all buyers excluded by player $i$ to pay for quantity $\bar{a}_i^j$. i.e.
$$
    c_i^j(s) = \displaystyle\int_0^{\bar{a}_i^j} P_i^j(z, s_{-i}) \ dz.
$$
This is the ``social opportunity cost" of the PSP pricing rule.

%The formulation is inspired to the thinnest allocation route for
%bandwidth given in \cite{lazar}. 
%We note that if a single seller $j$ can
%satisfy $i$'s demand, then
%(\ref{buyerutility}) reduces to the original form, defined in
%\cite{semret} as ``a simple buyer at a single resource element".

\subsubsection{Truthfulness.}
We will prove that the dominant strategy for buyers is to submit coordinated bids,
where all bids the buyer submits are equal. Our motivation for coordinated bids
comes from the idea of potential games. In potential games, the
incentive of all users to change strategy can be expressed as a single global
function. We map the incentive of a buyer over all auctions $j\in\mcI$ to a
single potential function. This is a standard method that is used often, as it
simplfies the analysis of both strategy and auction design. We will prove the necessary condition of an $\epsilon$-best
reply: a new bid price must differ from the last by at least $\epsilon$.
Thus, our strategic bid is an $\epsilon$-best response.

%\section{Strategic Framework}
\subsection{User Valuation (Strategic Incentive)}

We define a move to a better market position to be synonymous with a
strategic bid. \\ 
\textbf{Remark:} The terms ``bid" and ``strategy" are often interchangable,
from auction design and game theory, respectively.\\
Our mechanism allows a buyer
to \emph{opt-out} of auctions by submitting zero bids. This strategy maximizes
utility while minimizing the number of positive bids submitted to the overall
market. 
% OPT-OUT BUYER
We define each buyer as a user $i\in\mcI$ with quasi-linear utility
function $u_i = [u_i^j]_{j\in\mcI}$, a buyers' utility function is of the form,
\begin{equation}\label{buyerutility}
    u_i = \theta_i \circ \g_i(a) - c_i,
\end{equation}
where the composition of the elastic valuation function $\theta_i$ with $\g_i$
distributes a buyers' valuation of allocation $a$ across local markets (and thus
multiple sellers).
In this way we extend the PSP rules described
in \cite{semret} to design equilibria across subsets of local data-exchange markets.

The sellers, $j\in\mcI$ are not associated with an opt-out function. We
consider their valuation to be a functional extension of the buyers, where
$\theta^j$ is constructed from buyer demand.
The sellers' strategy can only be to determine the reserve price of their local
auction, using only information from buyers who have not opted out.
\\
\textbf{Remark:}
It is possible that a seller would be able to derive information about other
auctions by examining buyer bids over time, particularly if the seller had
knowledge of the buyer strategy. In this work, we assume sellers are unable to
derive opponent information from buyer bids. \\

Elastic valuation functions allow for even infinitesimal changes in the market
dynamics to be modeled. We give the definition for an elastic valuation
function as in \cite{lazar}.
\definition{(Elastic demand) \cite{lazar}}
A real valued function, $\theta(\cdot): [0,\infty) \rightarrow [0,\infty)$, is an \emph{(elastic) valuation
function} on $[0, D]$ if 
\begin{itemize}
    \item $\theta(0) = 0$,
    \item $\theta$ is differentiable,
    \item ${\theta}' \ge 0$, and ${\theta_i}'$ is non-increasing and continuous,
     \item There exists $\gamma > 0$, such that for all $z \in [0,D]$,
${\theta}'(z) > 0$ implies that for all $\eta \in [0, z), {\theta}'(z) \le
{\theta}'(\eta)
- \gamma(z - \eta)$. 
\end{itemize}

The elastic valuation of users and homogenous nature of data in the secondary market, 
allows for continuity of constraints imposed by the user strategies. 
We begin our analysis with buyer valuation $\theta_i$.
A buyers' valuation of an amount of data represents how much a buyer is willing
to pay for that amount. 
This is equivalent to the bid price, given a fixed amount of data,
satisfying $\theta_i$. We determine the
buyers' utility-maximizing bid given quantity $z\ge0$ to be a mapping to the lowest
possible unit price. We have,
% BUYER PRICING 
\begin{equation}\label{buyerinversedemand}
    f_i(z) \triangleq \inf\big\lbrace y\ge 0:
        \rho_i(y) \ge z, \ \forall \ j \in \mcI\big\rbrace,
\end{equation}
where $\rho_i(y)$ represents the demand function of buyer $i$ at bid price
$y\ge 0$, and gives the quantity that buyer $i$ would buy at a given price.
We determine that the market supply function corresponds to an extreme of
possible buyer demand, and 
acts as an ``inverse" function of $f_i$. We have, for bid price $y\ge 0$,
% BUYER DEMAND 
\begin{equation}\label{datasupply}
    \rho_i(y) = \sum_{j\in\mcI : p_i^j\ge y} D^j.
\end{equation}
We note that $f_i$ is such that
$i$ could still bid in \emph{any} auction $j\in\mcI$.
Therefore, the utility-maximizing bid price is the
lowest unit cost of the buyer to participate in all auctions, and corresponds to
the maximum reserve price amongst the sellers.

From the perspective of the
seller we have a more direct interpretation of valuation as revenue.
We determine the demand function of seller $j$ at reserve price $y\ge 0$ to be,
% BUYER DEMAND 
\begin{equation}\label{datademand}
    \rho^j(y) = \sum_{i\in\mcI : p_i^j\ge y} \g_i^j(a), 
\end{equation}
and define the ``inverse" of the buyer demand function for seller $j$ as
potential revenue at unit price $y$, we have, 
% SELLER REVENUE 
\begin{equation}\label{sellerinversedemand}
    f^j(z) \triangleq \sup\big\lbrace y\ge 0:
        \rho^j(y) \ge z, \ \forall \ i \in \mcI\big\rbrace,
\end{equation}
and, unsurprisingly, $f^j$ maps quantity $z$ to the highest possible unit data
price.

The valuation of any user must be modeled as a function of the entire
marketplace. 
Naturally, a buyers' valuation is aggragated over local markets, and the
sellers' valuation is aggragated over its own auction. We have already introduced the composition $\theta_i
\circ \g_i$ as the valuation of the buyers.
We further show that user valuation satisfies the conditions for an
elastic demand function, with valuations based on
(\ref{datademand}) and (\ref{sellerinversedemand}). 
We first note that, in
general (and so we omit the subscript/superscript notation), the valuation of data quantity $x\ge 0$ is given by,
$$
    \theta(x) = \int_0^x f(z) \ dz,
$$ 
as in \cite{semret}. Now, we have the following Lemma,
{
\lemma{(User valuation)}\label{uservaluation}
For any buyer $i\in\mcI$, the valuation of a potential
allocation $a$ is, 
\begin{equation}\label{buyervaluation}
    \theta_i \circ \g_i(a) = \displaystyle\sum_{j\in\mcI}
 \int_0^{\g_i^j(a)} f_i(z) \ dz.
\end{equation}
Now, we may define seller $j$'s valuation in terms of revenue,
\begin{equation}\label{sellervaluation}
    \theta^j= \displaystyle\sum_{i\in\mcI} \theta^j \circ \g_i^j(a) = \sum_{i\in\mcI}
 \int_0^{\g_i^j(a)} f^j(z) \ dz.
\end{equation}
We have that $\theta_i$ and $\theta^j$ are elastic valuation functions, with
derivatives
${\theta_i}$ and ${\theta^j}'$ satisfing the conditions of elastic demand.
}
\proof{
Let $\xi$
be a unit of data from buyer bid quantity $\g_i^j(a)$.
If $\xi$ decreases by incremental amount $x$, then seller bid $d_i^j$ must
similarly decrease. The lost potential revenue for seller $j$ is the price of
the unit times the quantity decreased, by definition, $f^j(\xi)x$, and so,
$$
    \theta^j(\xi) - \theta^j(\xi-x) = f^j(\xi)x.
$$
Thus (\ref{sellervaluation}) holds.
As we may use the same argument for (\ref{buyervaluation}), as such, we will
denote $f_i=f^j=f$ for the remainder of the proof. We observe that the function
$f$ is the first
derivative of the valuation function with respect to
quantity. Letting $\theta_i=\theta^j=\theta$, the existence of the derivative implies $\theta$ 
is continuous, and therefore, in this context, $f$ represents
the marginal valuation of the user, ${\theta}'$. Also, clearly
$\theta(0)=\theta(\sigma(0)) = 0$. Now, as we consider data to be an infinitely
divisible resource, we have a continuous interval between allocations $a$ and
$b$, where $a\le b$. Now, as $\theta$ is continuous, for some $c\in[a,b]$, 
$$
\theta'(c) = \lim_{x\to c}\frac{\theta(x) - \theta(c)}{x-c} = f(c),
$$
and so $f=\theta'$ is continuous at $c\in[a,b]$, and so as $a\ge 0$, 
$\theta'\ge 0$. Finally, we have that concavity follows from the demand
function. Then, as $\theta'$ is
non-increasing, we may denote its derivative $\gamma \le 0$,
and taking the derivative of the Taylor approximation, we have, $\theta'(z) \le
\theta'(\eta) + \gamma(z-\eta)$.
$\qed$}

The sellers' natural utility is the potential
profit, or simply $u^j = \theta^j$, where we have chosen to omit the original cost of the data
paid to the ISP, as it is not a component of our mechanism, and as a discussion
of mobile data plans is outside the scope of this paper.  
Now, a rational user will try
to maximize its utility, thus, user incentive manifests as a
response to market dynamics. A buyer has the choice to
opt-out of any auction, and as a seller will try to sell the maximum amount of
data, the highest possible reserve price is conditioned by ``natural" constraints. Utility-maximization 
acts as revenue maximization for a rational seller, and as cost minimization for a
rational buyer. Thus, for each user $p_i^j \ge \min\big(p_i^j)$ and $p_i^j \le
\max\big(p_i^j)$, which holds $\forall \ i, j\in\mcI$ such that $s_i^j>0$.
Now, rational buyer does not want to purchase extra data, as this would be
equivalent to overpaying, however $i$ submits positive bids to a set of
sellers, and a rational seller will attempt to maximize profit, and so will try and sell all of
its data. Therefore,
\begin{equation}\label{bidcap}
    \displaystyle\sum_{i\in\mcI} \g_i^j(a) \ge D^j \quad\text{and}\quad
\sum_{j\in\mcI} d^j_i \ge D_i,
\end{equation}
which holds $\forall \ i,j\in\mcI$. We will assume that buyers and sellers do not overbid, and so omit this
constraint from our formulation.
Thus, at equilibrium all users are satisfied, and 
$D^j = D_i$, although we observe that this result
does \emph{not} imply that $s_i=\nu^j$.

Finally, it is worth mention that the analysis of the auction as
a game assumes some forms of demand and supply, in order to derive properties.
The mechanism itself does not require any knowledge of user demand or
valuation.

\subsection{User Behavior}

Buyers and sellers are able
to change their bid strategies asynchronously and serially.
A users' local strategy space is therefore nondeterministic, and the prefererences of users
are subject to change, i.e. binary dependence.
Then, from \emph{Arrow's Theorem}, we have that no deterministic strategy can
provide a mapping of the preferences of users into a market-wide (complete
and transitive) strategy. As individual bids cannot map to a general objective, a better market
position can only be determined by an adaptive strategy.
We address the market risks and securities in our secondary data exchange market. 
We provide a game-theoretic model of a real market progression, which we use to
derive, then define, adaptive variables.

Assuming equal bandwidth for all users, we derive a globally optimal strategy suited for
users with local information in a distrubuted data-sharing model. 
In a multi-auction market, each auction a buyer joins has the potential to
decrease the potential cost of its data. However, increasing the size of the
auction implies a certain risk, which we may interpret as a potential and
definite
liability. Increasing the number of transactions causes additional messaging
overhead, fees, and increased competition from other buyers. 
A transaction also causes potential indirect costs, which may be considered work done to find
sellers, or effort of communication from participation. A seller
has the potential for greater profit with each new buyer in its auction, taking 
the same risk.
The liability of any user is naturally absorbed into the bid fee $\epsilon$,
as described in \cite{semret}. Therefore, according to our interpretation, the bid fee is
dependent on the association between two users and their market positions, in
addition to the underlying network structure. Now, both sellers and buyers must consider
the cost of adding additional users to their subsequent pools.

\subsubsection{Minimize crossover in buyer/ seller pools.}
Buyer $i$'s seller pool is determined by minimizing $n$, and is the smallest set of sellers that
allows for a coordinated bid, and the aggragate bids satisfy its demand, $D_i$.
\begin{align}
\begin{split}\label{buyercoordinate}
    \min\big\lbrace n \in \mcI \ &\vert \ n D^n\ge D_i \big\rbrace. 
\end{split}
\end{align}
Similarly, seller $j$ determines the minimal set of buyers that maximizes revenue and
sells all of its data, $D^j$.
\begin{equation}\label{sellercoordinate}
    \min\bigg\lbrace m \in\mcI \ \bigg\vert\displaystyle\sum_{i\in\mcI^j(m)}d_i^j
\ge D^j \bigg\rbrace,
\end{equation} 
We further determine that the set of buyers and sellers participating
in a single equilibrium is bounded by the potential indirect costs of
participation. We will denote this individual cost to each user as
$\varrho$. The indirect cost is the portion of the bid fee
$\epsilon$ that is dependent on the underlying network and the individual.
Observing that $\varrho$ indirectly effects user utility,
and therefore acts to establish a natural budget for each user. We define this
constraint as, 
\begin{equation}\label{poolbnd}
    u \le \varrho,
\end{equation}
which may be interpreted as the effort a rational user is willing to expend
on its message space, and serves to limit the size of the buyer/seller pools. 
This information may be collected from a specific
device's configuration, i.e. enabled roaming, daily data restrictions.
It is clear that an unconstrained
market, even with a finite number of users, could suffer from the
expense of many local auctions trading an infinitely divisible resource, thus
$\varrho$ is interpreted as
the ``liability" component of $\epsilon$ attempts to regulate network congestion.

\subsubsection{Buyer Strategy}

Although it is possible for a seller to fully satisfy a buyer $i$'s demand, it
is also reasonable to expect that a seller may come close to using their
entire data cap, and only sell the fractional overage. In this case, a buyer
must split its bid among multiple sellers. 
The buyer strategy bids in auctions with the highest quantities first,
a natural exploitation of the demand curve. A new seller entering the market
with a large quantity of data will be in high demand. This behavior contributes to market price
stability, as seller valuation is determined by buyer demand, the
buyer strategy tends towards equal valuation of all local markets, and therefore
similar prices. If a buyers' demand is not satisfied, they will need to bid in
markets with smaller data quantities, and so will bid on a larger portion of
the sellers' bid quantity, increasing their unit price. 
We define $j^* = n \le I$ represent the seller with the least amount of data
$\in\mcI_i$, i.e. $D^{j^*} \le D^j, \ \forall\ j \in\mcI^j$.
We define the composition,
$$
     \g_i^j \circ a = \g_i^j(a) = \frac{a_i^j}{j},
$$
to be the buyer strategy with respect to quantity.
We propose the following strategy. We include the proof in Appendix $A$.
% BUYER STRATEGY
{
\lemma{(Opt-out buyer strategy)}\label{buyerstrategy}
Let $i\in\mcI$ be a buyer and fix all other buyers' bids $s_{-i}$ at time
$t>0$, and let $a$ be $i$'s desired allocation. 
% OPT-OUT BUYER STRATEGY
Define,
\begin{equation}\label{opt-eq}
    \g_i^j(a) \triangleq \begin{cases} \g_i^{j^*}(a), \quad j\in\mcI^j,\\
            0, \quad\qquad j\ni\mcI^j.\end{cases}
\end{equation}
and bid price $p_i^j = \theta_i'(\g_i^j(a))$.
Now, (\ref{opt-eq}) holds $\forall \ j\in\mcI$.
}\\
Each time step, $\s^j$ is updated it is shared with all
participating buyers. At this point buyers have the opportunity to bid again,
where a buyer that does not bid again is assumed to hold the same bid, since a
buyer dropping out of the auction will set their bid to $s_i^j=(0,0)$.

% BUYER ALGORITHM
\begin{center}
\begin{algorithm}[H]
\caption{(Buyer response)}
\begin{algorithmic}[1]
\State $p_{i(0)} \gets \epsilon$, $s_{i(0)} \gets (p_i, D_i)$, $D_t \gets D_i$, compute
$\mcI_{i(0)}$
\State Update $s_i$ 
\While{$D_i(t) > 0$}
\State $D_{i(t+1)}^j \gets \sum_{j\in\mcI_i}\g_i^{j(t)}(a)$
\If{$D_{i(t+1)}^j < D_t$}
    \State Compute $\mcI_{i(t)}$
    \State $p_i \gets \theta_i(\g_i(a))$
\EndIf
\State $s_{i(t+1)} \gets (\g_i(a), p_i)$
\State Update $s_i$
\State $D_{i(t+1)}^j \gets D_{i(t)}^j$
\State $t \gets t+1$
\EndWhile
\end{algorithmic}
\end{algorithm}
\end{center}

Finally, we note that $\mcI_i$ is not the only
possible minimum subset $\in\mcI$ able to satisfy $i$'s demand, in fact, by
restricting the size of the set $\mcI_i$, we would be able to improve the
computation time of buyer $i$, at the cost of increasing the price.

\subsubsection{Individual rationality.}
We prove that a buyer cannot have a negative utility. Our
strategic framework creates an incentive for the seller to maintain a local
equilibrium, where supply equals demand. 
We define the reserve price for seller $j$ as,
\begin{equation}\label{newprice}
    p_*^{j} = p_{i^*}^{j} + \epsilon,
\end{equation}
where $i^*$ is the highest losing bidder with respect to bid price.
We claim that the choice of reserve price $p_*^j$ does not force any buyers out
of the local auction. 
A truthful bid implies that the new bid price differs from the last bid price by at least
$\epsilon$. As a seller must distribute bid vectors to all
buyers in its auction, we reason that the seller may employ a strategic caveat.
The seller will notify a buyer who is subject to a market shift by changing its
bid at the appropate index.

\subsubsection{Seller Strategy}
In order to
maximize revenue, the seller must also be able to respond dynamically to
strategic bids. In order to do this, we
determine that the seller may modify its reserve price in response to the
changing market dynamics.

Define any auction duration to be $\tau \in [0,\infty)$. 
We will show that sellers are able to maximize revenue in restricted subset of
buyers in $\mcI$, and as such will attempt to facilitate a local market
equilibrium for this subset. A local auction $j$ converges when $\forall \ i \in\mcI$,
$s_i^{j(t+1)} = s_i^{j(t)}$, at which point the allocation is stable, the data
is sold, and the auction ends. In the sellers' local environment, we determine
that the best course of action is to maximize revenue, and then try to keep its
buyer pool stable until convergence occurs. The exhaustive proof is included in
Appendix $B$.

% SELLER LOCAL STRATEGY
{
\lemma{(Localized seller strategy (i.e. progressive allocation))}\label{sellerstrategy}
For any seller $j$, fix all other bids $[s_i^k]_{i,k\ne j \in\mcI}$ at time $t>0\in\tau$.
For each $t\in\tau$, let $\omega(t)$ be given by (\ref{winner}), and perform
the update,
\begin{equation}\label{newdata}
    D^{j(t+1)} = D^{j(t)} - \g_{\omega(t)}^{j(t)}(a).
\end{equation}
Allowing $t$ to range over $\tau$, we have that $D^j=0$, and a local market equilibrium. 
}

Consider a user purchasing data from a subset of other network users.
The sellers' auction will function as follows: at each bid iteration all buyers
submit bids, and the winning bid is the buyer $i$ that has the highest price
$p_i^j$. The seller allocates
data to this winner, at which point all other buyers are able to bid again, and
the winner leaves the auction (with the exception where multiple bidders bid
the same price, where (\ref{dataallocation}) determines they will not fully
satisfy their demand, and so we will assume they remain in the auction). The auction progresses as such until all the
sellers' data has been allocated. 
% SELLER ALGORITHM
\begin{center}
\begin{algorithm}[H]
\caption{(Seller progressive allocation)}
\begin{algorithmic}[1]
\State $p^{j(0)} \gets \epsilon$, $s^{j(0)} \gets (p^j, D^j)$, $\bar\mcI =
\emptyset$, compute
$\mcI^{j(0)}$
\State Update $s^j$ 
\While{$D^j(t) > 0$}
\State $\bar{i} \gets \displaystyle\max_{i\in I^j}\sum_{i\in I^j} p_i^j$ 
\State $D^{j(t+1)} \gets D^{j(t)} - \g_{\bar{i}}^{j(t)}(a)$
%\State $p^j \gets \theta_{i^*}'(d_{i^*}^j)\circ e_i$
\State $p^j \gets p_{i^*}^j+\epsilon$ and $d^j \gets D^{j(t+1)}$
\State $s^{j(t+1)} \gets (d^j, p^j)$
\State Update $s^j$
\State $\bar\mcI \gets \bar\mcI \cup \bar{i}$
\For{$k \in \bar\mcI$}
\If{$p_k^j < p_{i*}^j$}
\State $D^{j(t+1)} = d_k^{j}$
\State $\bar\mcI \gets \bar\mcI \setminus \lbrace k \rbrace$
\EndIf
\EndFor
\State Compute $\mcI^{j(t)}$
\State $\mcI^{j(t+1)} = \mcI^{j(t)}\setminus \bar\mcI$
\State $t \gets t+1$
\EndWhile
\end{algorithmic}
\end{algorithm}
\end{center}

\subsubsection{Market Dynamics under Strategy}

We conclude this section by examining the relationship between the strategies of buyers and
sellers in local auctions. We model the impact of the dynamics of the
data-exchange market on a local auction $j$. As we have shown, the seller
is a functional extension of the buyer, with rules determined by the buyers'
behavior. This gives an auction $j$ a natural logical extension into the
global market through its buyers. We demonstrate that the symmetry between
buyer and seller behavior, consequently strategies, stretches into a symmetry across subsets of local
auctions. Additionally, we identify a clear bound restricting the influence of
local auctions on each other. Defining a single iteration of the auction, where a seller updates bid vector
$s^j$, and the buyers' response $s_i$, to comprise a single time step, and we
have the following Proposition,
% USER UNION
{
\proposition{(Valuation across local auctions)}\label{userunion}
For any $i,j\in \mcI$, 
\begin{equation}\label{jtoi}
j\in\mcI_i \Leftrightarrow i\in\mcI^j.
\end{equation}
Fix an auction $j\in\mcI$ with duration $\tau$ and define the influence sets of users. 
The primary influencing set is given as,
%(ALL THE SELLERS FROM EACH BUYERS POOL)
\begin{equation}\label{lambda}
    \Lambda = \displaystyle\bigcup_{i\in\mcI^j} \mcI_i,
\end{equation}
with secondary influencing set,
%(ALL THE BUYERS FROM EACH SELLER $i$ IN $j$'s BUYER POOL)
\begin{equation}\label{llambda}
    \lambda = \bigcup_{i\in\mcI^j} \bigg(\bigcup_{k\in\mcI_i} \mcI^k \bigg)
\end{equation}
\iffalse
(THE SELLERS ALL THE BUYERS IN $j$ HAVE IN COMMON! i.e. $j$ AND...)
\begin{equation}%\label{lambda}
    \Lambda = \displaystyle\bigcap_{i\in\mcI^j} \mcI_i,
\end{equation}
(THE COLLECTION OF BUYERS FROM EACH $i$ THAT ARE COMMON ACROSS ALL THE $i$'s SELLER
POOLS)
\begin{equation}%\label{llambda}
    \lambda = \bigcup_{i\in\mcI^j} \bigg(\bigcap_{k\in\mcI_i} \mcI^k \bigg)
\end{equation}
\fi
Define $\Delta = \Lambda \cup \lambda$.
Fixing all other bids $s_i^j \in \mcI$, and time $t>0\in\tau$, we have that,
\begin{equation}\label{union-eq}
    \displaystyle\sum_{j\in\Lambda} \theta_i^j = \sum_{i\in\lambda} \theta_i^j.
\end{equation}
}\\
\proof{
A local auction $j\in\mcI$, is determined by the collection of buyer bid
profiles, where buyer bid $s_i^j > 0 \Rightarrow j\in\mcI_i$. Using Proposition \ref{sellerstrategy} and (\ref{jtoi}), we have that, 
\begin{equation}\label{iff}
    i\in\mcI^j \Leftrightarrow p_i^j > p_{i^*}^j,
\end{equation}
where
(\ref{sellercoordinate}) defines $i^*$ as the losing buyer with the highest bid
price in auction $j$. By (\ref{buyerinversedemand}) $p_i^j \ge
p_{i^*}^j+\epsilon$, thus $p_i^j < p_{i^*}^j$ can only happen during a market
shift caused by the underlying dynamics.
Consider $k\in\mcI^j$ at time $t$ where, for example, some buyer(s) enter the auction, and
so (\ref{iff}) implies that
$\sum_{i\in\mcI^j} \g_i^j(a) > D^j$. Now, $p_i^j < p_{i^*}^j\Rightarrow k\ni\mcI^j$ and
$s_k^j>0$ will cause $k$ to initiate a shift. By Proposition \ref{buyerstrategy}, $k$ will set $s_k^j=0$, and begin to add sellers to its pool.
Suppose that at time $t$, $j$'s market is at equilibrium, i.e.
$\sum_{i\in\mcI^j} \g_i^j(a) = D^j$, and fixing all other bids, so 
no buyer $i\in\mcI^j$ rebids. Unless $k$ adds a seller with a higher reserve
price within $\vert\mcI^j\vert$ time steps, by (\ref{newdata}), $D^j=0$ and the
auction ends. Otherwise, at some time $t\in[t+1,\tau]$, we must have that
$\g_k^j \le D^j$, and $k$ rejoins auction $j$ or opts-out. Finally, overlooking market shifts and messaging
overhead, we have that,
    $\forall \ i\in\mcI^j$, $\nexists \ s_i^j > 0$ where $i \ni \mcI^j$, and (\ref{jtoi})
holds.
\iffalse
Furthermore, we note that $\lambda \subset \mcI$ is a complete and disjoint set
of bid vectors with respect to auction $j$. That is, $\forall \ i \in \mcI^j,
s_i^k > 0 \in \lambda$ if and only if $s_i^k>0 \in\mcI$, and so $s_i \in \lambda
\equiv s_i \in \mcI$. (NEED TO ADD TO LEMMA AND PROVE?)
\fi

Now, the subset $\mcI^j \subset \mcI$
determines $j$'s reserve price $p_{i^*}^j$. We will assume the buyer submits a
coordinated, truthful bid. Now, $\mcI_i\subset \mcI$ 
determines the unit price $p_{i}$ in buyer $i$'s bid.
The reserve price (\ref{newprice}) of seller $j$ is determined at each
shift, and is
the lowest price that $j$ will accept to perform any allocation.
Let $p_*^j = f^j \circ \g_i^j(a) $ denote the reserve
price of auction $j$, noting that $s_i^j=0, \ \forall \ i \in
[s_i^j]_{i\ni\mcI^j}$, and let $p_i^*= f_i \circ \g_i^j(a)$ denote the bid price of buyer $i$, i.e. $p_i^k = p_i^*,
\ \forall \ k \in\mcI_i$. 
Using Proposition
\ref{sellerstrategy}, for each $i \in \mcI^j$, we have from (\ref{buyerinversedemand}), (\ref{sellerinversedemand}), that $p_i^* \ge p_*^k,
\ \forall \ k \in\mcI_i$.

\iffalse
Now, we have,
\begin{equation}\label{rho}
    \rho^j(p_*^j) \le \displaystyle\sum_{i\in\mcI^j} e_i(a(s_i^j)),
\end{equation}
from (\ref{datademand}), (\ref{buyerinversedemand}),
(\ref{sellerinversedemand}), and so $p_i^j \ge p_{*}^j \ge p_i^k, \ \forall \ i \in
[p_i^j]_{i\in\mcI^j}$ and $\forall \ k \in [p_i^k]_{k\in\mcI_i}$
\fi

The incentive of each seller $\in\Lambda$ is to sell all of its data at the
best possible price. In the simplest case,
consider a disjoint local market $j$, where $\forall \ i\in\mcI^j$, $s_i^k = 0,
\ \forall \ k\ne j \in \mcI_i \Rightarrow \Lambda = \lbrace j \rbrace$ and
$\lambda = \mcI^j$. 
Again using (\ref{buyerinversedemand}) and  (\ref{sellerinversedemand}), it is
clear that $\theta_i = \theta^j, \ \forall \ i \in \mcI^j$.
In all other cases, the sellers $\in\Lambda$ are competing to sell
their respective resources to buyers whose valuations are distributed across multiple auctions. 
%By (\ref{jtoi}), $j\in \mcI_i, \ \forall \ i\in \mcI^j$. 
The set $\lambda$ represents all of the buyers influencing auction $j$, both
directly and indirectly. 
The bid price of buyer $i\in\mcI^j$ is determined by,
\begin{equation}\label{maxbid}
    p_i^* = \max_{k\in\mcI_i}\big(f^k \circ \g_i(a)\big)= \max_{k\in\mcI_i}(p_*^k).
\end{equation}
$\Lambda$ is the set of sellers directly influencing the bids of buyers
in auction $j$. Now, the reserve price for auction $j$ is such that,
\begin{equation}\label{resprice}
    p_*^j \le \min_{i\in\mcI^j}(p_i^*) - \epsilon,
\end{equation}
from (\ref{newprice}). Now, by Proposition \ref{sellerstrategy}, in the absence
of external influences caused by multi-auction market dynamics, we have 
that $j$ maintains a local market equilibrium from time $t$ to $(t+1)$.
From (\ref{lambda}) and (\ref{llambda}), $\Delta$ is defined by a seller $j\in\mcI$, where each
user $k\in\Delta$ has some direct or indirect influence on $j$. 
We may identify $\Delta$ by its dominant seller, and we denote $\Delta^j = \Lambda^j
\cup \lambda^j$.

Consider the set $\lambda^j$. For some buyer $i\in\mcI^j$, and then for
some seller $k \in \mcI_i$, we have a buyer $l \in \mcI^k$. By (\ref{jtoi}), $i,l \in \mcI^k$, and so the reserve price $p_*^k \le \min(p_l^*,
p_i^*)$, and $k, j\in \mcI_i\Rightarrow p_i^* \ge \max(p_*^k, p_*^j)$.
Suppose that $l \ni \mcI^j\Leftrightarrow j \ni \mcI_l$, so
that $p_l^* < p_*^j$, and
the valuation of buyer $l$ does not impact auction $j$ and vice versa, i.e. $\theta_l^j = 0$. Since $l\in\mcI^k$, $p_l^* \ge p_*^k\Rightarrow
p_*^k < p_*^j$, and $i\in\mcI^j \Rightarrow p_i^* \ge p_*^j$. Therefore, we
have that the ordering implied by (\ref{lambda}) and (\ref{llambda}) hold,
where,
\begin{equation}\label{ordering}
    p_*^k \le p_l^* < p_*^j \le p_i^*,
\end{equation} 
for any buyer $l\in\lambda^j$ such that $l\ni \mcI^j$. Now, suppose $\exists \ l\in\mcI^k$ such that $l\in\mcI^j \Rightarrow p_l^* \ge
p_*^j$. In the case where $p_l^* > p_i^*$, we must have that 
$\exists \ q \in \mcI_l$ such that $p_*^q > p_*^k$, % \Rightarrow \theta_l^q > \theta_l^k$. 
which implies, again by (\ref{maxbid}), $q\ni\mcI_i \Leftrightarrow i
\ni\mcI^q \Rightarrow p_*^q >
p_i^*$, therefore $\theta_i^q = 0$, and the reserve price of auction $q$ does not effect the valuation of
buyer $i$, and as $p_*^k < p_*^j \le p_i^* < p_l^*$, we examine $\mcI^j$ using
(\ref{iff}).
Lastly, in the case where $p_i^* > p_l^*$, by the same reasoning, $\theta_l^g =
0$, for some $g\in\mcI_i$.
We have that for any $l\in\mcI^k$ such that $l\ni
\mcI^j$, $\theta_l^j =0$, and when $l\in \mcI^j$, then either $\theta_i^q =0$,
where $q\in\mcI_l$, or $\theta_l^g = 0$, where $g\in\mcI_i$, and as $p_*^k <
p_*^j \le p_l^* < p_i^*$, we examine $\mcI^k$ using (\ref{iff}),
a shift in $\mcI^k$ causes a shift in $\mcI_i$, so that $\exists \ g\in\mcI_i$
such that $p_*^g \ge p_*^j$. Thus, we
determine a direct influence as $l\in\mcI^k
\cap \Lambda^j$, such that $p_l^* > p_i^*$,
and an indirect influence as, for any $l\in\mcI^k
\setminus \Lambda^j$, where $p_l^* > p_i^*$ results in $i^*\in\mcI^j$ initiating a shift.

Now, consider the subset $\Lambda^j$, by
Proposition \ref{userunion}, a shift occurs in 2 cases. (1) If $i\in\mcI^j$
decreases its bid quantity so that $\sum_{i\in\mcI^j}
\g_i^j(a) < D^j$, and (2) if buyer $i^*$, defined in Proposition
\ref{sellerstrategy}, increases its valuation so that $p_{i^*}^j <
p_*^j$. First, let buyer $i\in\mcI^j$ be the buyer in auction $j$ with the
lowest bid price, the ``lowest clearing player",
and further suppose $p_i^* > p_*^j + \epsilon$. That is, $\exists \ q \in \mcI_i$ such
that $p_*^q > p_*^j$. Fixing all other bids, a decrease in $q$'s demand will
directly impact buyer $i$. If at the end of the bid iteration, we still have that $i$ is the
buyer with the lowest bid price, then (\ref{sellerinversedemand}) holds and $j$'s valuation does not change. Otherwise a new $i^*$ will be chosen upon
recomputing $\mcI^j$, as in Proposition \ref{buyerstrategy}, and the market
will attempt to regain equilibrium.
Clearly, if $i^*$ in case (1) or resulting from case (2) increases in valuation,
then $p_*^j$ will similarly increase, by (\ref{uservaluation}). Consider the 
seller $k^*\in\mcI_{i^*}$ at time $t$, and suppose that $p_*^{k^*} \ge p_*^j$, however, we have that $p_{i^*}^j <
p_*^j \Rightarrow i^* \ni \mcI^j \Rightarrow k^*\ni\Lambda^j\Rightarrow \mcI^{k^*} \ni
\lambda^j$. Now, consider a buyer $l^*\in\mcI^{k^*}$. We need only consider the
case where $\exists \ k \in\Lambda^j$ such that $l^*\in\mcI^k\subset \lambda^j$
where we determine the influence of $\Delta^{k^*}$ on $\Delta^j$ by (\ref{iff}).

In each case we
have that (\ref{buyerinversedemand}) and (\ref{sellerinversedemand}) hold for
some fixed time $t$, and so, $\forall \ i \in\mcI^j$,
\begin{equation}\label{jdemand}
    \int_0^{\g_i^j(a)} f^j(z) \ dz 
    = \int_0^{\g_i^j(a)} f_i(z) \ dz,
\end{equation}
therefore $\theta_i = \theta^j, \ \forall \ i \in \mcI^j$.
Thus, any bid ouside of
our construction
has a zero valuation, with respect to buyers $\in\lambda$ and sellers
$\in\Lambda$, and therefore cannot cause shifts to occur except through a
shared buyer, e.g. some $l\in\mcI^k$. Thus, in all cases,
(\ref{buyerinversedemand}) and (\ref{sellerinversedemand}) hold. Fixing all
bids in any auction $q \ni \Lambda^j$,
we have, $\forall \ k\in\mcI_i$,
\begin{equation}\label{idemand}
    \displaystyle\int_0^{D^k} f^k(z) \ dz = \sum_{i\in\mcI^k}\int_0^{\g_i^k(a)}
f^k(z) \ dz,
\end{equation}
which holds $\forall \ k\in\mcI_i$, by (\ref{jtoi}) and Proposition
\ref{sellerstrategy}..
Finally, using (\ref{jdemand}), (\ref{idemand}), $\forall \ i\in\mcI^j, \forall \ k\in\mcI_i,\forall \ l\in\mcI^k$,
\begin{align}\label{tieqtk}
    \int_0^{\g_i^k(a)} f_i(z) \ dz &= \int_0^{\g_i^k(a)} f^k(z) \ dz,
\end{align}
and
\begin{align}\label{tkeqtl}
    \int_0^{\g_l^k(a)} f^k(z) \ dz &= \int_0^{\g_l^k(a)} f_l(z) \ dz.
\end{align}
Thus, with a slight abuse of notation for clarity,
\begin{equation}\label{funion-eq}
    \sum_{\lambda}\int_0^{\g(a)} f^\Lambda(z) \ dz =
\sum_{\Lambda}\int_0^{\g(a)} f_\lambda(z) \ dz,
\end{equation}
where the result follows by construction, and the continuity of $\theta'$.
$\qed$}

For completeness, in the case where the ISP $\kappa$ 
does not adhere to the market dynamics, so $p^\kappa > p^j +
\epsilon, \ \forall \ j \in\mcI$, then we may absorb the overage (difference) as part of the
bid fee.

The topology of the bids in $\delta$ is key to our result. We immediately have:

\subsubsection{Locally fair division.}
We claim that the allocation $a$ by seller $j$ for a local auction at
equilibrium is an \emph{equitable division}, a fair division where each buyer
equally values their valuation. We have that equitable division holds from
(\ref{tieqtk}) Proposition \ref{userunion}.

\subsubsection{Social welfare maximization.}
We define an optimal state of social welfare to be when valuations are equal
across a subset of local auctions. Then, $\Delta\subset \mcI$ to be a subset of
users where an optimal social welfare is achieved.

From these two properties, and Proposition \ref{userunion}, we have the
following Corollary, 
{
\corollary{($\Delta$-Pareto efficiency)}
    The subset $\Delta\subset \mcI$ is Pareto efficient, in that no user can
    make a strategic move without making any other user worse off.
}
\proof{Follows directly from the result of Proposision \ref{userunion}.}

% VCG ANALYSIS
\section{PSP Analysis}\label{analysis}

\subsection{Equilibrium}

We intend to show evidence shared network optima (a
global optimum). 
A buyer $i\in\mcI$ will have incentive to change its bid
quantity if it increases its opt-out value $\g_i$, and therefore its utility (\ref{buyerutility}). 
We will show that, without loss of utility, buyer $i$ may  
use a ``consistent" bid strategy within its seller pool, i.e. $d_i^j = d_i^k, \
\forall \ j,k \in \mcI_i$, and as such,
Proposition \ref{buyerstrategy} supports an optimal strategy with respect to
(\ref{buyerutility}). 
Our result shows that a buyer may select $\mcI_i$ in order to maximize its
utility while maintaining a coordinated bid strategy. 
Reasonably, if $j^*< I$, a buyer may increase the size of its seller pool
$\mcI_i$, thereby lowering its coordinated bid
quantity while obtaining the same (potential) allocation $a_i$. 
As buyer $i$ submits identical bids to multiple auctions, the bid price must be
as high as the highest reserve price $p_i^j \in \mcI_i$.
Buyer $i$'s bid then has identical bid price $p_i^j \forall \ j \in\mcI_i$. 
We further note that $i$ optimal strategy does not require reducing its bid price to a
minimum in each auction, where
the bid quantity $\g_i^j(a)$ is still fufilled. The pricing rule of the PSP
auction dictates that a buyer $i$ will pay the cost of excluding other players from
the auction, and as $i$'s bid price reflects its valuation of it data
requirement $D_i$ across all local markets, we have identical bid prices in
each auction where $s_i^j>0$. Obviously, if $j\ni\mcI_i$, then $\theta_i^j = 0$.

% BUYER COORDINATION
{
\lemma{(Opt-out buyer coordination)}\label{coordinationlemma} 
Let $i\in\mcI$ be a opt-out buyer and fix all sellers' profiles $s^j$.
For any profile $S_i = (D_i, P_i)$, let $a_i \equiv \sum_j a_i^j(s)$ be a
tentative data
allocation. For any fixed $S_{-i}$, a better reply for $i$ in any auction is $x_i =
\g_i \circ (z_i,y_i)$, where $\forall \ j \in \mcI_i$,
\begin{align*}
    z_i^j &= \g_i^{j}(a), \\
    y_i^j &= \theta_i'(z_i^j).
\end{align*}
Furthermore,
\begin{equation}\label{allocationcoordination}
    a_i^j(z_i,y_i) = z_i^{j},
\end{equation}
and
\begin{equation}\label{costcoordination}
    c_i^j(z_i,y_i) = y_i^{j},
\end{equation}
where $i$'s strategy is as in Proposition \ref{buyerstrategy}.
}\\
The proof is left to the Appendix, as it follows closely the work in
\cite{semret}.

\subsubsection{Incentive Compatibility}

The property of truthfulness is an essential component of equilibrium in
second-price markets. The strategies described in this paper have removed the
necessity for a user to determine its own valuation function, we intend to show that the market dynamics resulting from the
construction of the user strategy space results in truthful bids that are optimal for all users,
i.e. bid prices are to the marginal value as determined by market dynamics. 
To achieve incentive compatibility, we find that the opt-out buyer must choose
this subset so that its overall marginal value is greater than its market price.
We have so far only made the assumption of truthful bids throughout our analysis. As was shown in Lemma \ref{userunion}, a buyer only
has incentive to change its bid as a result of a market shift or partial
allocation. In a truthful reply, the term $\epsilon/\theta_i'(0)$
ensures that a new bid price differs from the last bid price by at least
$\epsilon$, thereby ensuring that a buyer does not change its bid without
correcting the effects of unstable shifts.
For a buyer $i$, define the set of possible $\epsilon$-best replies,
\begin{equation}\label{e-bid}
    S^\epsilon(s) = \lbrace s_i\in S_i(s_{-i}) : u(s_i;s_{-i}) \ge
u_i(s_i';s_{-i}) -\epsilon, \ \forall \ s_i'\in S_i(s_{-i})\rbrace,
\end{equation}
and the set of \emph{truthful} bids, 
\begin{equation}\label{t-bid}
    T_i = \lbrace s_i\in S_i(s_i) : z=\displaystyle\sum_{j\in\mcI_i} \g_i^j(a)
\wedge p_i = \theta_i'(z)\rbrace,
\end{equation}
where $\wedge$ denotes the logical ``and" operator. We note that the ``strategic" set $T_i$
is restricted by Proposition \ref{buyerstrategy}.
We have the following Proposition,

{
\proposition{(Incentive compatibility across local auctions)}\label{incentive}
Let $\Lambda$, $\lambda$ be defined as in Lemma (\ref{userunion}), and fix time
$t>0 \in \tau$, and fix $s^j, \ \forall \ j\in\Lambda$, and for some
buyer $i\in \mcI^j$, let $s_l$ also be fixed $\forall \ l\ni i \in \lambda$. Define,
\begin{align}
\begin{split}\label{eta}
    \chi_i &= \bigg\lbrace x\in [0,D_i] : 
 {\theta_i}'(x) > \max_{j\in\Lambda}P_i^j(x)\bigg\rbrace, 
\end{split}
\end{align}
and $z = \sup(\chi_i - \epsilon / \theta_i'(0))^+$, and for each $j \in
\Lambda$, 
$$
    v_i^j = \g_i^{j}(z),
$$
and 
$$
    w_i^j = \theta_i'(z).% \circ e = \theta_i'(v_i^j).
$$
Then a (coordinated) $\epsilon$-best reply for the opt-out buyer is $t_i =
(v_i,w_i) \in T_i\cap S_i^\epsilon(s_{-i})$, i.e., $\forall \ s_i, u_i(t_i;s_{-i}) + \epsilon \ge u_i(s_i;
s_{-i})$.
With reserve prices $p^j >0$, there exists a ``truthful" strategy game embeddeded
$\in\Delta$. Therefore, a fixed point $\in \Delta$ is a fixed point in
the multi-auction game. 
}\\
As in Lemma \ref{coordinationlemma}, we leave the proof to the Appendix.

The strategy space is comprised of a collection of bid, or ``strategy", vectors that
together, may be represented as a collection of potential functions, where change in buyer $i$'s
utility, resulting from a change in strategy, equals the change in
the local market objective of each seller $j\in\mcI_i$. These local objectives are known as potential functions, and are
formulated by mapping the incentives of all users in a local auction to a
single function. The goal of our analysis is to therefore construct a global potential function that encompasses
all local markets. Then, we may determine a Nash equilibrium by finding a local
optima of the potential function. Additionally, as the potential function also
iterates, it may be used in an analysis of convergence. The
convergence of a Nash equilibrium results from the progression of
$\epsilon$-best replies, where each subsequent bid is a unilateral
improvement, provided that $t_i$ is continuous in opponent profiles.
From the original proof by \cite{lazar}, we observe that the collection of
unconstrained truthful bids may be a subset
of the collection of $\epsilon$-best replies, i.e. $T_i\subset S_i^\epsilon$.         
%An essential component of determining the existence of a Nash equilibrium is 
For this work, it suffices to show the continuity of the set of truthful
$\epsilon$-best replies in the set of opponent bid profiles.
In order to address continuity in a global sense, we must demonstrate
continuity in the construction of our model.
Thus, we extend our analyis to be all-inclusive, %The game-theoritical basis of our argument is complemented by an abstract-theoretical analysis.
a.nd determine the
existence and ``uniqueness" of a global market objective by rigor of mathematical construction. Thus, we begin with the definition of
correspondence,
{
\definition{(Correspondence)}
A correspondence is mathematically defined as an ordered triple $(X,Y,R)$, where $R$ is a relation from $X$ to
$Y$, i.e. any subset of the Cartesian product $X\times Y$.
}

In an economic model, a correspondence $(S_i,S_{-i},R)$ defines a map from $S_i$
to the power set $S_{-i}$, where $R$ is a binary relation, i.e. $R \subset
S_i\times S_{-i}$. The classic
example of a correspondence in our model is the buyers' best response
$B_i^\epsilon$, where, for the multi-auction,
$S_i$ and $S_{-i}$ are built by repeatedly using the
cartesian product over bid profiles. The power set 
$S_{-i} =\Pi_j \big(\Pi_{k\ne i} S_i^j)$ arises naturally from the
product of ordered sets. 
The best response is a reaction correspondence
defined by the mixed-strategy game. Denoting  $B_i^\epsilon = T_i\cap
S_i^\epsilon$, is the set of truthful $\epsilon$-best replies in opponent bid
profiles $S_{-i}$. \\
\textbf{Remark:} 
The ease by which the game is constructed is a consequence of the the cartesian
product on a 2-dimensional message space. \\
A natural induced topology of this space is the product\footnote{This is known
as the Correspondence theory of truth}
topology, e.g. the canonical map $S_i \rightarrow \Pi_{j\in\mcI} S_i$.

Motivated by the symmetric nature of supply and demand, we determine the
game-theoritical argument is complemented by an abstract-theoretical analysis.
In fact, we may even be philosophically motivated, as the truth value of a bid
is determined only by how it relates to markets, and whether it provides an
accurate correspondence\footnote{This is known as the Correspondence theory of
truth}. Using a set-theoretical approach to address the sellers
bids, we derive our result from the symmetry of supply and demand,
(\ref{buyerinversedemand}) and (\ref{sellerinversedemand}), Proposition
\ref{incentive}, (\ref{newprice}), Lemma \ref{userunion} and Lemma
\ref{buyercoordinate}, and include the following corollary,
{
\corollary{Data-bid correspondence (seller cooperation)}
Let $\Delta$ be defined as in Lemma (\ref{userunion}).
For a fixed time $t\in (0,\tau]$, seller bid $s^j$ is consistent with a truthful
$\epsilon$-best reply.
}\\
\proof{
We claim there exists a binary equality relation $i\thicksim j$ that naturally evolves in the
strategy space. For a seller $j$, let $y=\theta_i'(\g_i^j(a))$ for a buyer $i$.
We use the the axiom of set equality,
based on first-order logic with equality, which states that,  
$\forall \ i\in\mcI, \ \forall \ j\in\mcI, (i\in \mcI^j \Leftrightarrow j\in\mcI_i)
\Rightarrow i \thicksim j$, and is a logical consequence of (\ref{jtoi}).
Then, for any allocation $a$, we may define the relation, $i\thicksim
j$,
\begin{equation}\label{eqrel}
        \big(\bar{D}_i^j(y), {\theta^j}'(\g_i^j(a))\big) =
        \big(\g_i^j(a),y)\big).
\end{equation}
Formally, the axiom states that a set is \emph{uniquely} determined by its
members. It follows that $\thicksim$ defines equality of bids using a static
analysis with respect to equilibrium, where all users who are not changing thier
bids are considered equal.\\
\textbf{Remark:} Equality is both an equivalence relation and a partial order,
and therefore is reflexive, transitive, symmetric and antisymmetric. \\
Now, we may define the mapping $s \mapsto [s]$,
\begin{equation}\label{indeqrel}
         1_\vartheta \equiv \theta_i'(z) - {\theta^j}'(z) > \epsilon,
\end{equation}o
noting that equality in the bid quantity is implicity satisfied and $z =
\bar{D}_i^j(y)\ge 0$. We have that $\vartheta$ is a price relation for a
buyer-seller pair. Let $S = \Pi_{j\in\mcI} \big(
\Pi_{i\in\mcI} S_i^j \big)$. The indicator function is the canonical mapping,
$1_\vartheta : S \rightarrow \lbrace{0,1\rbrace}$. Then, as the product
topology is preserved, the set of all indicator
functions on $S$ naturally forms the power set $\mathcal{P}(S) = S_i\times
S_{-i}$. Additionally, the set of all equivalence classes defines the quotient space, $S /
\thicksim \ \equiv \big\lbrace [k] : k\in\mcI \big\rbrace$, forming a partition
$P = \big\lbrace [s] : s\in S\big\rbrace$ of $S$. } Therefore each seller
$j\in\mcI$ is equivalent buyer(s) $i\in\mcI$. That is, buyers and
sellers with an equality relation (\ref{eqrel}) belong to the same local
auction equilbrium. Therefore, sellers' bids represent $\epsilon$-best replies
in the strategy space. $\qed$ (DONE? REALLY?)

{
\proposition{(Continuity of $\epsilon$-best reply on $\Delta$)}
Let $\Delta$ be defined as in Lemma (\ref{userunion}). For any buyer
$i\in\lambda^j$, the collection of bids $B_i$ is continuous in $S_{-i}$
}\\
\proof{
Define $\g_i \circ \overline{P} = \max_{i\in\mcI^j} \theta_i'(0)$, and
$\bar{P}_i(z, s_{i}) = \underline{P} = \epsilon -\varrho$, where $\epsilon$ is
the bid fee, and $\varrho$ is $i$'s liability estimate for auction $j\in\mcI$. 
We observe that $\g_i \circ B_i^\epsilon$ is simply $B_i^\epsilon$ restricted to seller pool
$\mcI_i$, i.e. $\g_i \circ B_i^\epsilon \equiv B_i^\epsilon\vert_{\mcI_i}$ . Thus,
we have $\g_i \circ T_i = ([0,D^k]_{k\in\mcI^j}\times [0,\g_i \circ
\overline{P}]^{\vert \mcI^j\vert})$
is a product of closed subsets of compact sets. Now, we have that a closed
subset of a compact set is compact and the resulting product topology gives
Tychonoff's theorem. every product of a
compact space is compact, we have $\g_i \circ B_i^\epsilon$
is compact subset of $B_i^\epsilon$.
 Now, letting $\overline{P} = \max_{i\in\lambda^j} \theta_i'(0)$, and we have by
definition of $\Delta$ and the product, 
\begin{align*}
    \g_i \circ& S_i(s_{-i}) \equiv g_i\vert_{\Lambda^j} : S_i \mapsto S_i \\
    &\Rightarrow\bigg(\bigcup_{i\in\mcI^j} [0,D^k]_{k\in\mcI_i},
[0,\overline{P}]\bigg)
    = \bigcup_{i\in\mcI^j}\bigg([0,D^k]_{k\in\mcI_i} \times
[0,\overline{P}]\bigg) \\
    &=  ([0,D^k]_{k\in\Lambda^j}, [0,\overline{P}]) \in \Lambda^j
\times \lambda^j \subset T.
\end{align*}
The result follows from the fact that $t_i$ is continuous in $s_i$, as was
proven in \cite{semret}, and as a finite union of compact sets is a compact set.
$\qed$}

As a result of user behavior, and subsequent strategies, we determine that the
data-exchange market behaves in a predictable way. However,
each auction may be played on the same or on a different scale in
valuation, time and quantity, and so the rate at which market fluctuations occur is
impossible to predict. 
We show that our bidding strategy results in (at least one), static Nash
equilibrium, where the sellers reserve prices are fixed. This result is
logically derived, and as our final result, we address the audience and note
that it is a mere step in the right direction.
{
\lemma{(Local Static Nash Equilibrium)} 
Let $\Delta$ be defined as in Lemma (\ref{userunion}), and suppose that auction
$j\in\Delta$ is at equilibrium. Fix all $s_i^j$. 
Using the rules of the data auction mechanism, along with type-based strategic
moves, $j$ converges to an $\epsilon$-Nash equilibrium. 
The proof follows closely that of \cite{semret}.
}\\
\proof{
As auction $j$ is at equilibrium, and 
since $\theta_i'$ is continuous, as was shown in Lemma \ref{uservaluation}, and
$t=[t_i^j] \in \lambda^j \times \Lambda^j$ is continuous in $s$ on
$T_k=\Pi_{k\in\Lambda^j} T_k^j$. Now, $t$
represents a continuous mapping of $[0,\sum_{k\in\Lambda^j} D^k]_{i\in\lambda^j}$ onto itself,
and we may use Brouwer's fixed point theorem, as in \cite{semret},
}

We have proven that buyers will submit bids according to their marginal
valuations. Now, in the interest of completeness (BAD SENTENCE), 
We have that all bids represent $\epsilon$-best replies, and, as was proven in
\cite{lazar}, the sellers' positive reserve price implies that bids are
truthful. Finally, by properties determined by the construction of a mixed strategy
symmetric game with a 2-dimensional message space, we may now restrict our
analysis to the set of continuous, truthful, $\epsilon$-best replies, $B^\epsilon$.

{
\corollary{Hemicontinuity of $\Delta$}
}

The data-sharing market consists of inter-dependent sets of these
multi-auction games around possible fixed points.  Clearly, the union of all possible
sets $\bigcup_{j\in\mcI} \Delta^j$ covers $\mcI$. We claim that the shared buyers
between the different subsets $\Delta$ form a sufficiently connected set that
the heirarchy described in Proposition \ref{userunion} holds. Then,
there can only be a single primary fixed point, where the sellers' reserve
price is an equilibrium price in the global market. 
We first address the analytical approach, and demonstrate properties of
$\Delta$ as a finite-dimensional linear topological space. We have that the
reserve price of the sellers, and the bid price of the buyers is constant
within an interval of length 2$\epsilon$. We have that 
(HERE) We have the following
Corollary, (MIGHT NEED TO REDEFINE.. NOT CLEAR IS A SEQUENCE, SHOULD BE
SEQUENCE OF PRICES INSTEAD OF USERS?) 
{
\corollary{(Primary fixed point)}
Let the set of shared buyers be denoted as,
$\underline{\lambda} = \bigcap_{j\in\mcI} \lambda^j$, and the set of all
sellers as, $\overline{\Lambda} = \bigcup_{j\in\mcI}\Lambda^j$. If
$\overline{\Lambda}$ is not a partition, i.e. $\nexists \ j,k
\in\overline{\Lambda}$ such that $\Lambda^j \cap \Lambda^k = \emptyset$, then,
for a fixed time $t$,
$\exists \ j \in\overline{\Lambda}$ such that $p_*^j \ge p_*^k, \ \forall \
k\ne j\in\overline{\Lambda}$. 
}\\
(CAN PROBABLY USE ALL BUYERS HERE... BECAUSE OF INF)
\proof{
We assume a finite number of users, with continuous valuation functions bounded both
above and below. From the assumption that $\overline{\Lambda}$ is not a
partition, we have that the limits exist with respect to bid price $p_i^j$, 
$$
    \limsup\limits_{j\to I} \overline{\Lambda} =  \bigcap_{j\ge 1}\bigcup_{k\ge j} \Lambda^k,
$$
is the primary seller $j$, and we have the market price $p_*^j$ from,
$$
    \liminf\limits_{i\to I} \overline{\lambda} = \bigcup_{i\ge 1}\bigcap{k\ge i}
\lambda^k,
$$
and the result follows from Lemma \ref{uservaluation}, Proposition
\ref{userunion} and Proposition \ref{incentive}.}
(NEED TO SHOW THEY ARE EQUAL...)
%$\qed$\\
(USE BOREL-CANTELLI WITH I.I.D? FUTURE WORK?)

We show that our bidding strategy is
part of a Nash equilibrium. We first show the existence of a \emph{static} Nash
equilibrium, where the sellers reserve prices are fixed.
{
\lemma{(Static Nash Equilibrium)} 
Let $\Delta$ be defined as in Lemma (\ref{userunion}), and
let the duration of auction $j$ be $\tau \in (0,\infty)$, and fix the
sellers reserve prices at $t\in(0,\tau), \ \forall \ j\in\mcI$.
Using the rules of the data auction mechanism applied independently by each
user, where users are acting according to their respective strategies, the multi-auction game converges to an $\epsilon$-Nash equilibrium. 
}\\
\proof{
(CAN'T DO THIS, NOT THE SAME TYPE OF GAME?)
As $\theta_i'$ is continuous, as was shown in Lemma \ref{uservaluation}, and
$t=[t_i^j] \in \lambda^j \times \Lambda^j$ is continuous in $s$ on
$T_k=\Pi_{k\in\Lambda^j} T_k^j$. Now, $t$
represents a continuous mapping of $[0,\sum_{k\in\Lambda^j} D^k]_{i\in\lambda^j}$ onto itself,
and we may use Brouwer's fixed point theorem, as in \cite{semret},
}

As a result of user behavior, and subsequent strategies, we determine that the
data-exchange market behaves in a predictable way. However,
each auction may be played on the same or on a different scale in
valuation, time and quantity, and so the rate at which market fluctuations occur is
impossible to predict (NEED HELP!). 
Arrow's paradox is an impossibility theorem stating that when buyers have three
or more distinct alternatives (auctions), no deterministic ranking system can
convert the ranked preferences of users into a market-wide (complete
and transitive) ranking while also meeting a specified set of criteria:
unrestricted domain, non-dictatorship, Pareto efficiency and independence of
irrelevant alternatives. It follows that the case where
$\theta_i = \theta^j$ as in (\ref{buyerinversedemand}) and
(\ref{sellerinversedemand}) will only occur if each set $\Lambda \cup \lambda$
is disjoint.

Nonetheless, we claim that our mechanism is normative, that irrelevant
alternatives should not matter, it is practical, uses minimal information,
strategy, and provides the right incentives for the truthful revelation of individual preferences.

The rules of the PSP
multi-auction drive market mutations that evolve and are regulated by the user
strategies.
(HERE)

(DEFINITION.. USE?)
In the General Symmetric Game, p is an evolutionarily stable mixed strategy if
there is a (small) positive number y such that when any other mixed strategy q
invades p at any level x < y, the fitness of an organism playing p is strictly
greater than the fitness of an organism playing q.
(EXPAND)
% DATA NASH EQ
{
\theorem{(Dynamic Nash Equilibrium)}
Using the rules of the data auction mechanism, the secondary market
converges to a $\epsilon$-Nash equilibrium. In the network auction
game with the data-PSP rules applied independently by each user according to their respective strategies, the secondary market converges to an $\epsilon$-Nash
equilibrium. 
}\\

\subsection{Efficiency}

Formally, the
mechanism is efficient, if, at equilibrium, the allocation maximizes
$\sum_i\theta_i(a_i)$.
(NEED OWN WORDS)
The objective in designing the auction is that, at equilbrium, resources al-
ways go to those who value them most. Indeed, the PSP mechanism does
have that property. This can be loosely argued as follows: for each player,
the marginal valuation is never greater than the bid price of any opponent
who is getting a non-zero allocation. Thus, whenever there is a player j
whose marginal valuation is less than player i 's and j is getting a non-zero
allocation, i can take some away from j , paying a price less than i 's marginal
valuation, i.e. increasing u i , but also increasing the total value, since i 's
marginal value is greater. Thus at equilibrium, i.e. when no one can unilat-
erally increase P their utility, the total value is maximized.

\subsection{Convergence}

% CONCLUSION AND FUTURE WORK
\section{Conclusion and Future Work}

We point out the need for better management of data on the consumer
level. It is obvious that there is profit to be made by
supplying data to the data-driven consumer. However, consumers must be given a
choice. Consider Bob, a mobile wireless network user. Marketing firms have already realized that an incentive
for Bob to interact with their ads is to provide them at no cost to his mobile
    data. Inevitebly, ISPs will find a way to supply the growing demand. We
    have demonstrated the trend towards a Monopoly of knowledge\footnote{John
    Watson (2006), suggests that monopolies of knowledge gradually suppress new ways
    of thinking.}, and suggest
    that trust, and policy be decentralized, for the greater good.

Mathematically, we have shown that if truthfulness holds locally for both buyers and sellers, i.e.
$p_i ={\theta_i}', \ \forall \ j \in \mcI_i$ and $p^j = {\theta^j}', \ \forall \
i \in\mcI^j$, then, in the absence of market shifts,
there exists an $\epsilon$-Nash equilibrium extending over a subset of connected
local markets. Observing the symmetric, natural topology of the strategy space,
we conjecture that a unique subspace limit exists for connected $\Delta$. A
study of this metric is the direction of our future work.

Finally, we note the harmonious relationship of users, data, and mathematics.
The exchange of data in the secondary market must mimic the topology of the
strategy space. Thus, although we assume that the message space of PSP auctions
is too small, we conjecture that this relationship is the backbone of an
evolutionary system, naturally derived from the continuity of demand and
shifting market equilibria.

\appendix

\section{Strategy}
\subsection{Proof of Lemma 3.5}
\proof{
We assume that a buyer will try and fill their data requirement.
In the case that there exists a seller who can completely satisfy a buyers'
demand, $j^*=1$, $\vert \mcI_i\vert =1$ and (\ref{buyercoordinate}) holds. If such a buyer does not exist,
as the set $\mcI_i$ is ordered by the quantity of the sellers' bids, $i$ may discover 
$j^*$ by computing $\mcI_i$.
Suppose that $D_i >
\sum_{j\in\mcI}D^j$, then $j^* > I$ and $\mcI_i = \emptyset$. 
We model the ISP at time $t>0$ as a seller $\kappa$ with bid $s^\kappa
=(d^\kappa, p^\kappa)$,
where $d^\kappa > D^j, \
\forall \ j\in\mcI_i$, and
$p^\kappa$ represents the price for data set by the ISP, which we note is also
the upper bound of the sellers' pricing function. We note that in \cite{zheng}
this cost is the data overage fee.
Consider some $k \ne i \in \mcI$ where $p_i^j = p_k^j$. 
The allocation rule (\ref{dataallocation}) determines that the data will be
split proportionally
between all buyers with the same unit price. 
It is possible that the resulting partial allocation of data to $i$ and $k$
would not satisfy some demand. As the two cases $i$ and $k$ are the same, we
will only consider one. Suppose seller $j$ updates its
bid to reflect the new data quantity, where
$d_i^{j(t+1)} < \g_i^{j(t)}(a)$. First, $i$ sets its bid to $s_i^j = 0$, and
from the new subset $\mcI_i$, submits bids until
$\sum_{j\in\mcI_i} \g(a)_i^j \ge D_i$, by (\ref{bidcap}).
Now, we consider the case where a new buyer $k$ with bid
price $p_k^j > p_i^j$ for some $j \ \in \mcI_i$, in other
words, a new buyer $k$ may enter the market with a better price, decreasing the
value of $i$'s bid for $j\in\mcI_i$. In this case, by (\ref{buyercoordinate}),
$i$ will choose $\mcI_i$ so that, $\g_i^{j(t+1)}(a) = \g_i^{j(t)}(a) - \g_k^{j(t)}(a)$,
and so $\mcI_i$ is large enough to balance the additional demand from
$k$. Finally, we consider the case where $\vert\mcI^j\vert = I$, where the
demand of buyer $i$ exceeds the supply, and the case where $\g_i(\varrho) >
\theta_i(\g_i(a))$, where the overhead exceeds the current valuation of the
data. Then, by (\ref{buyerinversedemand}), the valuation of the data increases until either
the demand is satisfied, the debit from the overhead costs are balanced
(\ref{poolbnd}), or the upper bound of the sellers' reserve price
$p^\kappa$ is reached. Thus, as in each case we have that $i$ is able to satisfy thier demand,
and we determine that the opt-out strategy is optimal.
$\qed$}

\subsection{Proof of Lemma 3.6}
\proof{
We assume that the seller will try to maximize its revenue. In the case where
$\vert\mcI^j\vert =1$, then if $\g_i^j(a)=D^j$, then $j$'s market is at equilibrium.
Otherwise, we arrive at the case of multiple buyers, which we note includes the
case where $\g_i^j(a) < D^j$, which is reflected trivially here.

For auction $j$ with multiple buyers, $i^*$ is the \emph{losing} buyer with the highest unit
price offer, determined by (\ref{sellercoordinate}), 
Suppose that for some $i\in\mcI^j$, buyer demand is not met. 
In this case, by (\ref{bidcap}) the seller must notify $i$ of a
partial allocation by changing the bid vector at index $i$.
With this caveat, and Proposition \ref{buyerstrategy}, we have that the
aggragate demand of subset $\mcI^j$ is satisfied by seller $j$.
Although the buyers' valuation $\theta_i$ is not known to the seller, we will assume
that buyers are bidding truthfully, and so the new reserve price $p_{i^*}^j +
\epsilon ={\theta_{i^*}}' + \epsilon$. For clarity, let the reserve price be denoted by
$p_*^j$.
Now, by the elasticity of (\ref{buyerinversedemand}) and
(\ref{sellerinversedemand}), we have that, $\forall \ z\ge 0$,
$f_{i^*}(z) < f^j(z) \le f_i(z)$,
which holds $\forall \ i \in \mcI^j$, and $\forall \ j \in \mcI_i$.
We claim that the choice of reserve price $p_*^j$ does not force any buyers out
of the local auction. To show this, we use the assumption of truthful bids, and
the fact that since the auction begins at time $t>0$, buyers will
bid at least once. As will be addressed in further analysis, we assume
that a new bid price differs from the last bid price by at least
$\epsilon$. Suppose the auction starts at equilibrium, so $\sum_{i\in\mcI^j}
\g_i^j(a) = D^j$ at time $t=0$. The reserve price $p_*^j$ set at time $t=0$
begins the auction with the first bid iteration, and so at $t>0$, $\forall \ i \in
\mcI^j$, we have that $p_i^j - p_*^j \ge \epsilon$.
Now, in the case where at $t=0$, $\sum_{i\in\mcI^j} \g_i^j(a) > D^j$, by
(\ref{dataallocation}), the seller notifies (any) buyer $k$ with the lowest bid
price of a partial allocation by changing
$d_k^j$ thus by Proposition \ref{buyerstrategy}, $k$ either decreases its
demand or increases its valuation until
$\g_i^j(a) \le d_i^j$. Then, as the seller computes the set $\mcI^j$ at each time
step, a new $i^*$ may be chosen and the buyers bid again.
Suppose $\exists \ k \in \mcI^j$ such that $\forall \ l \ \in
\mcI_k, \ i\ni \mcI^l \ \forall \ i\ne k\in\mcI^j$. That is, $k$ is
disconnected from all other buyers $i\in\mcI^j$, and suppose that $d_k^j$
is partial allocation at $t>0$, and further suppose that there are many
$l\in\mcI_k$ where $\vert \mcI^l\vert > \vert\mcI^j\vert$. The more buyers an
auction has, the more likely that cases will occur that cause buyers to rebid,
particularly if auctions $l\in\mcI_k$ have overlapping buyers, then $k$ may
opt-out of auction $j$, i.e. $s_k^{j(t)} \ne s_k^{j(t+1)} =
0$, then the seller may simply return the tentatively allocated data to $D^j$.
Finally, we note that if for some $i\in\mcI^j \ \exists \ k\in\mcI^j$
such that $p_{i}^j = p_k^j$, then the seller again notifies the buyers of a
partial
allocation by changing $d_i^j$ and $d_k^j$ by (\ref{dataallocation}).
Thus we determine the valuation between seller
$j$ and buyer $i$ is well-posed, the reserve price (\ref{newprice}) is
justified, and we have a local equlibirium at time $\tau$.
$\qed$}

\section{Equilibrium}
\subsection{Proof of Lemma 4.1}
\proof{
As $s_{-i}$ is fixed, we omit it, in addition, we will use $u\equiv u_i \equiv u_i(s_i) \equiv
u_i(s_i;s_{-i})$. In full notation, we intend to show
$$
    u_i((d_i,p_i);s) \le u_i((z_i,y_i);s_{-i}).
$$
Now, if there exists a seller who can fully satisfy $i$'s demand,
then $\vert\mcI_i\vert = 1$, and the case is trivial as no
coordination is necessary for a single bid. %(SPECIAL CASE OF MONOPOLY? THINK!)\\
Otherwise, buyer $i$'s demand can only be satisfied by purchasing data from
multiple sellers.
We will show that $i$ may increase $\vert\mcI_i\vert$, and so decreasing
$d_i^j, \ \forall \ j\in\mcI_i$, without decreasing $\sum_{j\in\mcI_i}u_i^j$.
Buyer $i$ maintains ordered set $\mcI_i$ where the sellers with the
largest bid quantities are considered first; the index of seller $j^*$ defines
a minimal subset $\mcI_i$, satisfying (\ref{buyercoordinate}). By construction, 
$d_i^{j^*}$ is the minimum quantity bid offered by any $j\in \mcI_i$.
Thus by (\ref{buyercoordinate}) and (\ref{opt-eq}), $\forall \ j\in \mcI_i$,
$k\ni\mcI_i$, $\g_i^k(a) \le z_i^j = \g_i^{j}(a)$,
and so, using (\ref{union-eq}), 
\begin{equation}\label{minimaleq}
    \g_i^{j}(a)  \le \bigg\lbrack D^{j} - \sum_{k\in \mcI^j : p_k^{j}> y_{i}^{j}} d_k^{j}\bigg\rbrack^+.
\end{equation}
Now, the buyer valuation function (\ref{buyervaluation}), guarantees
that $\forall \ j\in\mcI_i$, $y_i^j \ge p_{i^*}^j$, where $p_{i^*}^j$ is the reserve price of seller $j$, defined in Proposision
\ref{sellerstrategy}, and is by definition the minimum price for a buyer bid to
be accepted. 
As $\bar{D}_i^j$ is non-decreasing, $\forall \ j\in \mcI_i$,
$k\ni\mcI_i$,
$$
    D_i^j(y_i^j) \ge D_i^j(p_i^{j})\ge D_i^j(p_i^{k}).
$$
\iffalse
Furthermore, suppose there exists buyer $k\in \mcI^j$, such that
$\g_k^{j}(a) \ge \g_i^{j}(a)$ and $D^j - \sum_{k\in\mcI^j,k\ne i}d_k^j < z_i^{j}$, resulting in a
partial allocation. Bid $s_k^j \in s_{-i}$, and so is considered in buyer $i$'s strategy. 
As buyer $i$ is allowed to choose subset $\mcI_i$, and $\mcI_i$ is a minimal
set, (\ref{buyercoordinate}) states that $n$ is such that $\g_i^{j}(a)
\ge \g_k^{j}(a)$ for any $k\in\mcI$, therefore such a buyer $k$ cannot exist.
We have, $\forall \ j\in \mcI_i$, $k\ni\mcI_i$,
$$
    {\theta_i}'(\g_i^{k}(a)) \ge {\theta_i}'(\g_i^{j}(a)) \ge y_i^j.
$$
Now, suppose that $p_k^j = p_i^j$, however $d_k^j > d_i^j$...
\fi
Thus (\ref{minimaleq}) holds and so, by (\ref{dataallocation}),
\begin{align*}
     a_i^j(z_i,p_i) &=\min_{i\in\mcI^j}\bigg(z_i^j, \bigg\lbrack D^{j} - \sum_{p_k^{j}> y_i^{j}}
d_k^{j}\bigg\rbrack^+\bigg) \\
    &= z_i^{j}= \g_i^{j}(a)
\end{align*}
where the last equality is by definition, and so (\ref{allocationcoordination}) is proven.
From (\ref{datacomposed}), $\bar{D}_i^j(y, s_{-i}) =
0 \ \forall \ y < p_{i^*}^j$, and $\bar{D}_i^j(y, s_{-i}) =0 \le \epsilon
\Rightarrow \g_i^j(a) = 0 \Rightarrow z_i^k = 0, \ \forall \ k \ni \mcI_i$, and therefore,
\begin{align*}
    \displaystyle\sum_{j\in\mcI_i} c_i^j(z_i,y_i) =\sum_{j\in\mcI_i}
c_i^j(z_i,p_i),
\end{align*}
thus (\ref{costcoordination}) simply shows that changing the price $p^j_i$ to $y^j_i$ does not exclude any
additional buyers, as the bid $p^j_i$ was already above the reserve price of any
seller $j \in \mcI_i$.
We proceed to show that $x_i$ does not result in a loss of utility for
buyer $i$, that is,
$$
    u_i \le u_i(z_i,y_i).
$$
From (\ref{allocationcoordination}), we have $a_i^{j}(z_i,y_i) = z_i^{j} = \g_i^j(a(z_i,y_i))$, 
and so,
$$
    \theta_i\circ \g_i^j(a(z_i,y_i)) = \theta_i\circ \g_i^j(a),
$$ 
which holds $\forall \ j\in\mcI_i$.
Therefore, by the definition of utility (\ref{buyerutility}),
and the buyers' valuation (\ref{buyervaluation}), 
\begin{align*}
    &\theta_i\circ \g_i(a(z_i,y_i)) - \theta_i(a)\circ \g_i(a) \\
    &= u_i(z_i,y_i) - u_i =  \displaystyle\sum_{j\in\mcI_i} c_i^j -c_i^{j}(z_i,y_i)\\
    &= \sum_{j\in\mcI_i}\int_{a_i^j(z_i,p_i)}^{a_i^j} f_i(d_i^j - x) \ dx.
\end{align*}
Then, as $a_i(z_i,p_i) \le z_i^j \le a_i^j $, and noting that
$z_i^j>0\Rightarrow \theta_i\ge 0\Rightarrow f_i\ge 0$, we have $u_i(z_i,y_i) - u_i \ge 0$, $\forall \ j\in\mcI_i$.
$\qed$}
%Finally, we note that $z_i^j$ does not increase the aggragate demand of the buyer.

\subsection{Proof of Proposition 4.2}
\proof{
We claim that $t_i$ is an $\epsilon$-best reply for buyer $i$. That is,
$$
    u_i(t_i;s_{-i}) + \epsilon \ge u_i(s_i;s_{-i}).
$$
As a result of auction initialization, a seller $j$'s valuation defines its
reserve price to be determined
by a buyer $i \ni \lambda$, even if this price is zero, we have that $p^j =
\epsilon \ge 0 \ \forall \ j \in\Lambda$.
Let $z = \sup(\chi_i^j)$, and again let $p_*^j = f^j \circ \g_i^j(a) $ denote the reserve
price of auction $j$, and $p_i^*= f_i \circ \g_i^j(a)$ denote the (coordinated) bid price of buyer $i$. 
We have that $i\in\mcI^j$, 
and (\ref{buyerinversedemand}) defines $\theta_i'(z)$ as being
max of the reserve prices $p_*^j,\ \forall \ j\in\mcI_i$, therefore
(\ref{eta}) is such that,
$$
    \theta_i'(z) > \max_{j\in\Lambda}P_i^j(v_i^j),
$$ 
which implies, as $\theta_i'$ is non-increasing and $P_i^j \ge 0$,
we have $\forall \ j \in \mcI_i$, 
\begin{align*}
    w_i^j &> P_i^j(v_i^j) \\
    \Rightarrow v_i^j &\le D_i^j(w_i^j) = D^j - \rho^j(w_i^j).
\end{align*}
And so, by (\ref{dataallocation}),
\begin{align*}
    a_i^j(t_i;s_{-i}) &= v_i^j \\
    \Rightarrow \displaystyle\sum_{j\in\Lambda} a_i^j(t_i;s_{-i}) &= z.
\end{align*}
Therefore, $\forall \ j\in\Lambda$ and $\forall \ i \in \lambda$ such that
(\ref{tieqtk}) and (\ref{tkeqtl}) hold,
\begin{align*}
    \int_0^{v_i^j} \bar{P}_i(x) \ dx 
    = \sum_{j\in\Lambda}\int_0^{\g_i^j(z)} P_i^j(x) \ dx.
\end{align*}
It follows that,
$$
    u_i(t_i;s_{-i}) = \int_0^{z} \theta_i'(x) \ dx -  \g_i \ \circ \int_0^{z} \bar{P}_i(x) \ dx.
$$
Suppose $\exists \ s_i = (d_i, p_i)$ such that $u_i^j(s_i;s_{-i}) > u_i^j(t_i;
s_{-i}) + \epsilon$. Propositions
\ref{coordinationlemma} and
\ref{buyerstrategy}, define the coordinated bid, $\nu_i=(\zeta_i, p_i)$,
% = (\g_i\circ a_i(s), \ \theta_i'(z)) = (z_i, c_i),$
using (\ref{tieqtk}) and (\ref{tkeqtl}),
for each $j\in \Lambda$, $\g_i^j(a_i^j(\nu_i; s_{-i})) = \zeta_i^j$,
then clearly $u_i(\nu_i, s_{-i}) \ge u_i(s_i, s_{-i}) \Rightarrow u_i(t_i;s_{-i}) -
u_i(s_i;s_{-i}) > \epsilon$. Denoting $\zeta_i^j$ (fixed) as $\zeta$, 
$$
    \int_z^{\zeta} {\theta_i}'(x) \ dx-\int_z^{\zeta} \bar{P}_i(x) \ dx > \epsilon.
$$
For concave valuation functions, the first-order derivative of $\theta$ at point
$0$ gives the maximum slope of the valuation function, and so the factor $\epsilon
/ \theta'(0)$ guarantees that new bids will differ by at least
$\epsilon$, and as such, buyer $i$ will remain in
any local auction with reserve price determined by (\ref{newprice}). We therefore
verify that,
$$
    \displaystyle\int_{z}^{z +\epsilon / \theta_i'(0)} \theta_i'(x) \
dx \le \epsilon,
$$
and as $P_i^j \ge 0$, we have that, from the construction of $\zeta$,
$$
    \displaystyle\int_{z+\epsilon / \theta_i'(0)}^\zeta \theta_i'(x) \ dx -
\int_{z+\epsilon / \theta_i'(0)}^\zeta \bar{P}_i(x) \ dx > 0.
$$
If $\zeta > z + \epsilon / \theta_i'(0)$, then for some $\delta >0$,
$\theta_i(z+\epsilon / \theta_i'(0)+\delta) > P_i^j(z+\epsilon /
\theta_i'(0)+\delta)$, contradicting (\ref{eta}).
Now, if $\zeta \le z$, then $\theta_i'(z+\epsilon / \theta_i'(0)) < P_i^j(z+\epsilon /
\theta_i'(0))$, also a contradiction of (\ref{eta}), and so buyer $s_i$ cannot
exist.
Finally, as we may consider $\Delta \subset \mcI$ to be a
multi-auction game, our user strategies form a ``truthful" local game 
with strategy space restricted to $\epsilon$-best replies from buyers $\in
\lambda$. Therefore we have that a fixed point in the ``truthful" game is a fixed point for the auction. 
$\qed$}

\section{A Simple Example}
\example
Finally, we give a simple example of convergence to a local market equilibrium,
where the buyers are assumed to respond with their truthful,
$\epsilon$-best replies.

\begin{center}
\begin{tabular}{c|c|c}
Name & Bid total & Unit price\\
\hline
A & 50 & 1\\ 
B & 40 & 1.2\\
C & 26 & 1.5\\
D & 20 & 2\\
E & 14 & 2.2
\end{tabular}
\end{center}
Let $s^{(1)} = [(65,\epsilon)]_{i\in\mcI}$ and
$s^{(2)}=[(85,\epsilon)]_{i\in\mcI}$. The buyer bids are as follows: 
\begin{align*}
    s_A &= [(0,0), (50,1)], \\
    s_B &= [(0,0), (40,1.2)], \\
    s_C &= [(0,0),(26,1.5)], \\
    s_D &= [(0,0),(20,2)], \\
    s_E &= [(0,0), (14,2.2)]. 
\end{align*}
Then at $t=1$, we have 
bid vector $s^{(2)}=[(0,p^{(2)}), (20,p^{(2)}), (26,p^{(2)}),\\ (20,p^{(2)}),
(14,p^{(2)})]$, and so  $(D^{(2)},p^{(2)}) = (85, 1+\epsilon)$, The buyer response is,
\begin{align*}
    s_A &= [(50,1), (0,0)], \\
    s_B &= [(40,1.2), (0,0)], \\
    s_C &= [(0,0),(26,p^{(2)})], \\
    s_D &= [(0,0),(20,p^{(2)})], \\
    s_E &= [(0,0),(14,p^{(2)})].
\end{align*}
At $t=2$, $(D^{(1)},p^{(1)}) = (65,1+\epsilon)$, with
bid vector $s^{(1)}=[(25,p^{(1)}),\\ (40,p^{(1)}), (0,0), (0,0),
(0,0)]$. $(D^{(2)},p^{(2)}) = (25, 1+\epsilon)$. Then,
\begin{align*}
    s_A &= [(25,p^{(1)}), (25,p^{(2)})], \\
    s_B &= [(40,p^{(1)}), (0,0)], \\
\end{align*}
where we have removed bids to indicate winner(s) with a tentative allocation.
At $t=3$, $(D^{(1)},p^{(1)}) = (50,1+\epsilon)$, with
bid vector $s^{(1)}=[(25,p^{(1)}), (40,p^{(1)}), (0,0), (0,0),
(0,0)]$. $(D^{(2)},p^{(2)}) = (0, 1+\epsilon)$ and $s^{(2)}=[(25,p^{(1)}),
(0,0), (26,p^{(2)}), (20,p^{(2)}),\\(14,p^{(2)})]$. Then,
\begin{align*}
    s_A &= [(25,p^{(1)}), (0,0))].
\end{align*}
At $t=4$ the auction ends.\\
\textbf{Remark:} In the case where market resources do not satisfy (\ref{bidcap}), however as
this constraint is not restricted in time, we reason that in the case of
insufficient data in the market buyers may wait for additional sellers or
purchase from the ISP, $\kappa$, as a monopoly sale. Similarly, in
the case of insufficient demand, where we may assume that data is held at time
$t=0$ by $\kappa$ at bid price $\epsilon$.

%\end{multicols}